---
title: "Testing analysis methods"
author: "Darwin Del Castillo"
format: 
  docx:
    number-sections: true
editor: source
---

# Generating simulated data
Data generated in Stata using methods in [Methods for dealing with time-dependent confounding](https://onlinelibrary.wiley.com/doi/10.1002/sim.5686).

TMLE methods tested with `ltmle` package. A comprenhensive review of the methods and their implementation in R is available in [Using longitudinal targeted maximum likelihood
estimation in complex settings with dynamic interventions](doi.org/10.1002/sim.8340) and [Targeted Maximum Likelihood Estimation for Causal Inference in Observational Studies] (doi.org/10.1093/aje/kww165).

```{r simulate}
#| output: false
here::i_am("report/01_basic_analysis.qmd")

if (!requireNamespace("RStata", quietly = TRUE)) {
  message("RStata not installed, installing RStata package...")
  install.packages("RStata")
}
pacman::p_load(RStata)
stata_path <- "/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp"
options("RStata.StataPath" = stata_path)
options("RStata.StataVersion" = 18.5)

stata(here::here("code","simulated_data","sim_data.do"))
```

```{r read_simulated}
#| output: false

df1 <- haven::read_dta(here::here("data","sim_data","sim1.dta"))
df2 <- haven::read_dta(here::here("data","sim_data","sim2.dta"))
df3 <- haven::read_dta(here::here("data","sim_data","sim3.dta"))
```

# Causal DAG
The DAG for the simulated data DF1 is as follows:
```{r dag1}
#| output: false
#| include: true
#| fig-cap: "DAG for simulated data with single time point treatment"
#| fig-width: 8
#| fig-height: 6

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=LR]

                  node[fontname=Arial]
                  L1 [label=<<B><I>L1</I></B>>]
                  A0 [label=<<B><I>A0</I></B>>]
                  A1 [label=<<B><I>A1</I></B>>]
                  Y  [label=<<B><I>Y</I></B>>]
                  U0 [label=<<B><I>U0</I></B>>,fontcolor=deepskyblue4,style=dashed,color=deepskyblue4]

                  A0->L1->A1->Y
                  A0->Y
                  A0->A1
                  U0->A0 [style=dashed, color=deepskyblue4]
                  U0->L1 [style=dashed, color=deepskyblue4]

                  {rank=same; A0}
                  {rank=same; L1; A1}
                  {rank=max; Y}
                  {rank=min; U0}
                  }
                  "
)
```


The DAG for the simulated data DF3 is as follows:
```{r dag}
#| output: false
#| include: true
#| fig-cap: "DAG for simulated data"
#| fig-width: 8
#| fig-height: 6

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=LR]

                  node[fontname=Arial]
                  L0 [label=<<B><I>L<SUB>0</SUB></I></B>>]
                  L1 [label=<<B><I>L<SUB>1</SUB></I></B>>]
                  L2 [label=<<B><I>L<SUB>2</SUB></I></B>>]
                  L3 [label=<<B><I>L<SUB>3</SUB></I></B>>]
                  A0 [label=<<B><I>A<SUB>0</SUB></I></B>>]
                  A1 [label=<<B><I>A<SUB>1</SUB></I></B>>]
                  A2 [label=<<B><I>A<SUB>2</SUB></I></B>>]
                  A3 [label=<<B><I>A<SUB>3</SUB></I></B>>]
                  Y  [label=<<B><I>Y</I></B>>]

                  L0->L1->L2->L3
                  A0->A1->A2->A3

                  L0->A0
                  L1->A1
                  L2->A2
                  L3->A3

                  A0->L1
                  A1->L2
                  A2->L3

                  L0->Y
                  L1->Y
                  L2->Y
                  L3->Y
                  A0->Y
                  A1->Y
                  A2->Y
                  A3->Y

                  {rank=same; L0; A0}
                  {rank=same; L1; A1}
                  {rank=same; L2; A2}
                  {rank=same; L3; A3}
                  {rank=max; Y}
                  }
                  "
                  )
```

```{r prep_analysis}
#| output: false

# Loading helpers
source(here::here("fnct", "testing_helpers.R"))

# Loading packages
pacman::p_load(tidyverse,
               gfoRmula,
               gFormulaMI)

# Reshaping data
names(df1) # single-time point treatment, binary A
names(df2) # single-time point continuous treatment, continuous A
names(df3) # time-varying confounder and treatment

dt3 <- df3 |> 
  rename(y3 = y) |>
  mutate(id = row_number()) |> 
  select(-u) |> 
  pivot_longer(l0:y3, names_to = c("var", "t0"), values_to = "val", names_pattern = "(.)(.)") |>
  pivot_wider(names_from = var, values_from = val) |> 
  mutate(t0 = as.integer(t0)) |> 
  data.table::as.data.table()
```

# Potential outcomes for all treatment combinations
The potential outcomes for all treatment combinations are as follows, according to the paper:


|  a0 |  a1 |  a2 |  a3 | True mean w/PO |
|----:|----:|----:|----:|---------------:|
|   0 |   0 |   0 |   0 |            2.1 |
|   0 |   0 |   0 |   1 |            1.6 |
|   0 |   0 |   1 |   0 |            1.6 |
|   0 |   0 |   1 |   1 |            1.1 |
|   0 |   1 |   0 |   0 |            1.6 |
|   0 |   1 |   0 |   1 |            1.1 |
|   0 |   1 |   1 |   0 |            1.1 |
|   0 |   1 |   1 |   1 |            0.6 |
|   1 |   0 |   0 |   0 |            1.6 |
|   1 |   0 |   0 |   1 |            1.1 |
|   1 |   0 |   1 |   0 |            1.1 |
|   1 |   0 |   1 |   1 |            0.6 |
|   1 |   1 |   0 |   0 |            1.1 |
|   1 |   1 |   0 |   1 |            0.6 |
|   1 |   1 |   1 |   0 |            0.6 |
|   1 |   1 |   1 |   1 |            0.1 |

# Testing with `gfoRmula`
Testing trajectories for binary treatment with time-varying confounder using `gfoRmula` on `df3`.
```{r gformula-test}
time_points <- 4
# static <- 1L  # gfoRmula intervention-type code for "static"

# Complex interventions
intvars <-  as.list(rep("a", 16))

interventions <- tibble(a = 0:1, b = 0:1, c = 0:1, d = 0:1) |>
  expand.grid() |>
  arrange(a, b, c, d) |>
  t() |>
  as_tibble() |>
  unclass() |>
  map(~list(c(static, .x)))

int_description <-
  map(interventions, ~paste(.x[[1]][2:5], collapse = "-") |> paste0("Exposed: ", x = _))


gf_out <- gformula(
  dt3,
  outcome_type = "continuous_eof", # eofu
  id = "id", # i(i)
  outcome_name = "y", # out(y)
  covnames = c("l", "a"), # com(y:regress, l:logit, a:logit)
  covtypes = c("binary", "binary"),  # see above
  covparams = list(
    covmodels = c(
      l ~ lag1_a*lag1_l,
      a ~ l*lag1_a
    )
  ),
  time_name = "t0", #t(t)
  time_points = time_points,
  intvars = intvars,
  interventions = interventions,
  int_descript = int_description,
  histories = c(lagged),
  histvars = list(c("a", "l")),
  ymodel = y ~ a + l + lag1_a + lag2_a + lag3_a + lag1_l + lag2_l + lag3_l ,
  seed = 1234,
  # parallel = TRUE,
  # nsamples = 100, # Number of bootstrap samples 
  nsimul = 2000
  # ncores = 6
)

gf_out$result |> 
  filter(Interv. != 0) |> 
  transmute(Intervention = int_description, Estimate = round(`g-form mean`, 3)) |> 
  knitr::kable()
```

# Testing with `IPTW- MSM`
## Case 1: Single time point treatment, binary treatment
```{r msm-iptw-test-single-point}
# Step 1: marginal treatment probability
mod_a0 <- lm(a0 ~ 1, data = df1)
p0 <- ifelse(df1$a0 == 1, fitted(mod_a0), 1 - fitted(mod_a0))

# Step 2: P(A1 = observed | A0, L) — conditional on prior treatment and confounder
mod_a1 <- lm(a1 ~ a0 * l, data = df1)
p1 <- ifelse(df1$a1 == 1, fitted(mod_a1), 1 - fitted(mod_a1))

# Step 3: Unstabilized IPTW weights
w <- 1 / (p0 * p1)

# Step 3.1: Stabilized IPTW weights
# Numerator: P(A1 = observed | A0 only)
mod_a1_num <- lm(a1 ~ a0, data = df1)
p1_num <- ifelse(df1$a1 == 1, fitted(mod_a1_num), 1 - fitted(mod_a1_num))

w_s <- (p0 * p1_num) / (p0 * p1)

# Step 4: MSM outcome model
msm_fit <- lm(y ~ a0 * a1, data = df1, weights = w)

msm_fit_s <- lm(y ~ a0 * a1, data = df1, weights = w_s)

summary(msm_fit)
summary(msm_fit_s)
```

## Case 2: Single time point treatment, continuous treatment
```{r msm-iptw-test-continuous-l}
# Denominator: f(A1 | A0, L) — conditional density
mod_a1_d <- lm(a1 ~ a0 * l, data = df2)
sigma_d   <- sigma(mod_a1_d)
f_denom   <- dnorm((df2$a1 - fitted(mod_a1_d)) / sigma_d)

# Numerator: f(A1 | A0) — no L (stabilized)
mod_a1_n <- lm(a1 ~ a0, data = df2)
sigma_n   <- sigma(mod_a1_n)
f_num     <- dnorm((df2$a1 - fitted(mod_a1_n)) / sigma_n)

# Stabilized weights
w_s <- f_num / f_denom

# MSM outcome model
msm_fit_cont   <- lm(y ~ a0 * a1, data = df2, weights = w_s)
summary(msm_fit_cont)
```

## Case 3: Time-varying treatment and confounder
Testing trajectories for binary treatment with time-varying confounder using MSM with IPTW
```{r msm-iptw-test-time-varying}
# --- Denominator models: P(At = obs | past treatment, current L) ---
d0 <- glm(a0 ~ l0,      family = binomial, data = df3)
d1 <- glm(a1 ~ a0 * l1, family = binomial, data = df3)
d2 <- glm(a2 ~ a1 * l2, family = binomial, data = df3)
d3 <- glm(a3 ~ a2 * l3, family = binomial, data = df3)

pd0 <- ifelse(df3$a0 == 1, fitted(d0), 1 - fitted(d0))
pd1 <- ifelse(df3$a1 == 1, fitted(d1), 1 - fitted(d1))
pd2 <- ifelse(df3$a2 == 1, fitted(d2), 1 - fitted(d2))
pd3 <- ifelse(df3$a3 == 1, fitted(d3), 1 - fitted(d3))

# --- Numerator models: P(At = obs | past treatment only) ---
n0 <- glm(a0 ~ 1,  family = binomial, data = df3)
n1 <- glm(a1 ~ a0, family = binomial, data = df3)
n2 <- glm(a2 ~ a1, family = binomial, data = df3)
n3 <- glm(a3 ~ a2, family = binomial, data = df3)

pn0 <- ifelse(df3$a0 == 1, fitted(n0), 1 - fitted(n0))
pn1 <- ifelse(df3$a1 == 1, fitted(n1), 1 - fitted(n1))
pn2 <- ifelse(df3$a2 == 1, fitted(n2), 1 - fitted(n2))
pn3 <- ifelse(df3$a3 == 1, fitted(n3), 1 - fitted(n3))

# --- Unstabilized and stabilized weights ---
w   <- 1 / (pd0 * pd1 * pd2 * pd3)
w_s <- (pn0 * pn1 * pn2 * pn3) / (pd0 * pd1 * pd2 * pd3)

summary(w_s)  # mean ≈ 1; flag if max >> 20
summary(w)    # mean >> 1; flag if max >> 20

# --- MSM outcome model ---

msm_tv <- lm(y ~ a0 + a1 + a2 + a3, data = df3, weights = w_s)
summary(msm_tv)
```

```{r tmle-test}
pacman::p_load(ltmle, SuperLearner, xgboost)

# Two reasons SL.xgboost fails inside ltmle:
# 1. Binary A/L nodes are passed as R factors (encoded 1/2, not 0/1).
# 2. The Q-model for continuous bounded Y uses family=binomial (TMLE targeting
#    step), so xgboost receives non-0/1 labels with binary:logistic objective.
# Both cases: drop to gaussian so xgboost uses reg:squarederror.
SL.xgboost.ltmle <- function(Y, X, newX, family, ...) {
  if (is.factor(Y)) Y <- as.integer(Y) - 1L
  if (!all(Y %in% c(0, 1))) family <- gaussian()
  SL.xgboost(Y = Y, X = X, newX = newX, family = family, ...)
}

# ltmle requires Y in [0,1] — scale and back-transform after
y_range   <- range(df3$y)
df3_ltmle <- df3 |>
  select(l0, a0, l1, a1, l2, a2, l3, a3, y) |>
  mutate(y = (y - y_range[1]) / diff(y_range))

# --- ATE: always treated vs never treated ---
ltmle_ate <- ltmle(
  data          = df3_ltmle,
  Anodes        = c("a0", "a1", "a2", "a3"),
  Lnodes        = c("l0", "l1", "l2", "l3"),
  Ynodes        = "y",
  abar          = list(treat = rep(1, 4), control = rep(0, 4)),
  SL.library    = "glm",
  estimate.time = FALSE
)
summary(ltmle_ate)

# --- All 16 regimes (matches g-formula table) ---
regimes       <- expand.grid(a0 = 0:1, a1 = 0:1, a2 = 0:1, a3 = 0:1) |> arrange(a0, a1, a2, a3)
regime_labels <- apply(regimes, 1, paste, collapse = "-")
regime_list   <- split(regimes, seq_len(nrow(regimes)))

# Helper: run ltmle for all 16 regimes with a given SL library
run_ltmle <- function(sl_lib) {
  map(regime_list, ~ltmle(
    data          = df3_ltmle,
    Anodes        = c("a0", "a1", "a2", "a3"),
    Lnodes        = c("l0", "l1", "l2", "l3"),
    Ynodes        = "y",
    abar          = as.numeric(.x),
    SL.library    = sl_lib,
    estimate.time = FALSE
  ))
}

ltmle_glm <- run_ltmle("SL.glm")
ltmle_xgb <- run_ltmle(c("SL.glm", "SL.xgboost.ltmle"))

# Extract back-transformed estimates for a list of ltmle results
extract_est <- function(results) {
  map_dbl(results, ~summary(.x)$treatment$estimate * diff(y_range) + y_range[1])
}

tibble(
  Regime  = regime_labels,
  GLM     = extract_est(ltmle_glm),
  XGBoost = extract_est(ltmle_xgb)
  ) |>
  knitr::kable(digits = 3)
```