---
title: "New pipelines: multiple imputation"
author: "Darwin Del Castillo"
format: 
  docx:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from waves 3 and 10, which were collected from 2011 to 2020. 

```{r setup}
#| output: false
here::i_am("report/05_imputation.qmd")

pacman::p_load(gfoRmula,
               gFormulaMI,
               DiagrammeR,
               tidyverse,
               purrr,
               magrittr,
               colorBlindness,
               ltmle,
               mice,
               mori,
               SuperLearner,
               targets,
               knitr)

source(here::here("fnct", "import_cleaning.R"))
source(here::here("fnct", "helpers.R"))
```

# Model specification: Two waves of data, time-invariant covariates + PCS
In this model, we include 3 waves of data (waves 3 to 5) and adjust for time-invariant covariates and the physical component summary (PCS) score. The model is specified as follows: \
**Notation** \
- $TI_{cov}$: Time-invariant covariates: \
    - Sex \
    - Ethnicity \
    - Highest educational qualification \

```{r model1}
#| label: model1
#| echo: false
#| include: true
#| fig-cap: "Estimated mean MCS scores at wave 5, adjusting for time-invariant covariates and PCS score. Nodes are summarised by time point to avoid clutter."
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=TB]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  C1 [label=<<B><I>PCS<SUB>t-1</SUB></I></B>>]
                  C2 [label=<<B><I>PCS<SUB>t</SUB></I></B>>]
                  C3 [label=<<B><I>PCS<SUB>t+1</SUB></I></B>>]
                  E1 [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E2 [label=<<B><I>FD<SUB>t+1</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E3 [label=<<B><I>FD<SUB>t+2</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y2 [label=<<B><I>MCS<SUB>t+1</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y3 [label=<<B><I>MCS<SUB>t+2</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E1 -> E2 -> E3
                  Y1 -> Y2 -> Y3
                  C1 -> C2 -> C3

                  TI -> E1, E2, E3, C1, C2, C3, Y1, Y2, Y3

                  C2 -> E2
                  C2 -> Y2

                  C3 -> E3
                  C3 -> Y3

                  E1 -> C2
                  E2 -> C3

                  C1 -> E1
                  C1 -> Y1

                  E1 -> Y1 
                  E2 -> Y2
                  E3 -> Y3

                  Y1 -> C2
                  Y2 -> C3

                  Y1 -> E2
                  Y2 -> E3

                  {rank=min; TI}
                  {rank=same; C1, C2, C3}
                  {rank=same; E1, E2, E3}
                  {rank=max; Y1, Y2, Y3}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point. Heavy compute (cleaning, reshaping, mice imputation, LTMLE fitting, gFormula and IPTW sensitivity analyses) is orchestrated by `_targets.R` at the repo root; this report consumes the cached outputs.

```{r data_prep}
#| output: false
wide_data <- tar_read(wide_data)
seed_random <- 20260522
```

# Multiple imputation (wide format)
```{r mice-gformula}
#| output: false
wide_mids <- tar_read(wide_mids)
```

## Additional: mice diagnostics
```{r mice-diag}
#| fig-width: 12
#| fig-height: 14
#| fig-align: center
#| output: false
#| include: false
#| eval: false

# Diagnostic plots for mice imputation
bwplot(wide_mids)

## Checking goodness of fitness
fit <- with(wide_mids, glm(ici(wide_mids) ~ sf12mcs_dv_0 + pcs_lagged_0 + econ_dist_bin_0 + sf12mcs_dv_1 + pcs_lagged_1 + econ_dist_bin_1 + sf12mcs_dv_2 + pcs_lagged_2 + econ_dist_bin_2 + hiqual_dv_base + race_base + sex_dv_base,
                          family = binomial))

ps <- rep(rowMeans(sapply(fit$analyses, fitted.values)),
          wide_mids$m + 1)

numeric_vars <- colnames(wide_data_mi)[
  vapply(wide_data_mi, is.numeric, logical(1))
]

fitted_plot <- xyplot(
  wide_mids,
  as.formula(paste(paste(numeric_vars, collapse = " + "),
                   "~ ps | as.factor(.imp)")),
  outer    = TRUE,
  as.table = TRUE,
  xlab     = "Probability that record is incomplete",
  ylab     = "Value",
  pch      = c(1, 19),
  cex      = 1.1,
  col      = mdc(1:2),
  scales   = list(
    y = list(relation = "free", cex = 0.7),
    x = list(cex = 0.7)
  ),
  par.strip.text = list(cex = 0.7)
)

fitted_plot

# Convergence plot
plot(wide_mids)

# Worm plot for one outcome
wp_mids <- function(mids_obj, var, level = 0.95) {
  long <- mice::complete(mids_obj, action = "long", include = TRUE)
  long$.imp <- factor(long$.imp,
                      levels = 0:mids_obj$m,
                      labels = c("Observed",
                                 paste("Imp.", seq_len(mids_obj$m))))
  long$.z <- as.numeric(stats::ave(long[[var]], long$.imp,
                                   FUN = function(v) scale(v)[, 1]))
  long <- long[!is.na(long$.z), ]

  crit <- qnorm((1 + level) / 2)

  lattice::qqmath(
    ~ .z | .imp, data = long,
    aspect   = 1,
    as.table = TRUE,
    xlim     = c(-3, 3),
    ylim     = c(-0.6, 0.6),
    xlab     = "Unit normal quantile",
    ylab     = "Deviation",
    main     = sprintf(var),
    par.strip.text = list(cex = 0.7),
    panel = function(x, ...) {
      qq    <- stats::qqnorm(x, plot.it = FALSE)
      detr  <- qq$y - qq$x
      n     <- length(qq$y)
      z_seq <- seq(-2.75, 2.75, by = 0.05)
      p     <- pnorm(z_seq)
      se    <- sqrt(p * (1 - p) / n) / dnorm(z_seq)

      lattice::panel.abline(h = 0, lty = 2, col = "grey60")
      lattice::panel.lines(z_seq,  crit * se, lty = 2, col = "grey40")
      lattice::panel.lines(z_seq, -crit * se, lty = 2, col = "grey40")
      lattice::panel.points(qq$x, detr,
                            pch = 19, cex = 0.4, col = mice::mdc(2))
    }
  )
}

wp_mids(wide_mids, "sf12mcs_dv_0")
wp_mids(wide_mids, "sf12mcs_dv_1")
wp_mids(wide_mids, "sf12mcs_dv_2")
wp_mids(wide_mids, "pcs_lagged_0")
wp_mids(wide_mids, "pcs_lagged_1")
wp_mids(wide_mids, "pcs_lagged_2")

# density plots for imputed vs observed data
densityplot(wide_mids, ~ sf12mcs_dv_0)
densityplot(wide_mids, ~ sf12mcs_dv_1)
densityplot(wide_mids, ~ sf12mcs_dv_2)

densityplot(wide_mids, ~ pcs_lagged_0)
densityplot(wide_mids, ~ pcs_lagged_1)
densityplot(wide_mids, ~ pcs_lagged_2)

densityplot(wide_mids, ~ econ_dist_bin_0)
densityplot(wide_mids, ~ econ_dist_bin_1)
densityplot(wide_mids, ~ econ_dist_bin_2)

densityplot(wide_mids, ~ hiqual_dv_base)
```

# `mice` positivity check
```{r mice-positivity}
#| output: false
#| include: true
#| eval: false

positivity_results <- function(mids_obj) {
  mice::complete(mids_obj, action = "long", include = FALSE) |>
    group_by(.imp,
             econ_dist_bin_0,
             econ_dist_bin_1,
             econ_dist_bin_2) |>
    summarise(n = n(), .groups = "drop") |>
    group_by(econ_dist_bin_0,
             econ_dist_bin_1,
             econ_dist_bin_2) |>
    summarise(
      mean_n = round(mean(n), 0),
      min_n  = min(n),
      max_n  = max(n),
      n_imp_with_zero = sum(n == 0),
      .groups = "drop"
    )
}

print(positivity_results(wide_mids))
```

# Main analysis: LTMLE with multiple imputation
The LTMLE sweep over the 8 regime × imputation pairs is computed by the targets DAG (`R/fit_ltmle_one.R` + `R/pool_ltmle.R`); each branch dispatches as a separate SLURM job via `crew.cluster`. The pooled tibble is read in below.

```{r ltmle-mi}
#| output: false
#| include: true
ltmle_results <- tar_read(ltmle_results)
```


# Sensitivity analyses: g-formula and IPTW

## G-formula via multiple imputation

### Imputing counterfactual outcomes under each regime using `gFormulaMI`
```{r gformula-mi}
#| include: false
#| output: false
mi_results <- tar_read(mi_results)
```

## IPTW with multiple imputation

```{r iptw-mi}
#| output: false
#| include: false
iptw_fit  <- tar_read(iptw_fit)
iptw_pool <- mice::pool(iptw_fit)
summary(iptw_pool)
```

### Table with results from all models
```{r table-results}
#| output: true
#| include: true
iptw_results <- tar_read(iptw_results)
comparison   <- tar_read(comparison)

comparison |> knitr::kable(digits = 3)
```

```{r plot_results}
#| output: true
#| include: true

# Combine results into a single data frame for plotting
comparison |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("ltmle", "mi", "iptw_sw"),
                    labels = c("LTMLE", "G-Formula MI", "IPTW")),
    intervention = factor(intervention,
                          levels = c("0-0-0", "0-0-1", "0-1-0", "0-1-1", "1-0-0", "1-0-1", "1-1-0", "1-1-1"),
  )) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Intervention",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(39, 51, by = 2)) +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

ggsave(
  filename = here::here("figs", "results_comparison.png"),
  plot     = results_comparison,
  width = 6, height = 8, dpi = 300
)

```