---
title: "Models with full set of covariates (time-invariant + PCS)"
author: "Darwin Del Castillo"
format: 
  html:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from waves 3 and 10, which were collected from 2011 to 2020. 

```{r setup}
#| output: false
here::i_am("06_full_models.qmd")

pacman::p_load(gfoRmula,
               gFormulaMI,
               DiagrammeR,
               tidyverse,
               purrr,
               magrittr,
               colorBlindness,
               ltmle,
               mice,
               SuperLearner,
               targets,
               knitr)

# Quarto renders from report/, but the targets store lives at the project root. 
# Passing absolute path to every tar_read() below so resolution doesn't depend on the working directory.
tar_store <- here::here("_targets")

source(here::here("fnct", "import_cleaning.R"))
source(here::here("fnct", "helpers.R"))
```

# Model specification: Two waves of data, time-invariant covariates + PCS
In this model, we include 3 waves of data (waves 3 to 5) and adjust for time-invariant covariates and the physical component summary (PCS) score. The model is specified as follows: \
**Notation** \
- $TI_{cov}$: Time-invariant covariates: \
    - Sex \
    - Ethnicity \
    - Highest educational qualification \
    - Region of residence (GOR) \
- $PCS_{t-1}$, $PCS_{t}$, $PCS_{t+1}$: Physical component summary score at time points t-1, t, and t+1 \
- $IC_{t}$, $IC_{t+1}$, $IC_{t+2}$: Intermediate confounders (lagged) at time points t, t+1, and t+2, which include: \
   - Number of dependents (0, 1, 2+) \
   - Economic benefits receipt (yes/no) \
   - Home ownership status (yes/no) \
   - Marital status (partnered vs not partnered) \
- **Time-varying confounder:** age

```{r model1}
#| echo: false
#| include: true
#| fig-cap: "Direct Acyclic Graph showing causal assumptions for the present analysis. Only two nodes per variable are presented to avoid clutter."
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=TB]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  A1 [label=<<B><I>Age<SUB>t</SUB></I></B>>]
                  A2 [label=<<B><I>Age<SUB>t+1</SUB></I></B>>]
                  C1 [label=<<B><I>CF<SUB>t-1</SUB></I></B>>]
                  C2 [label=<<B><I>CF<SUB>t</SUB></I></B>>]
                  I1 [label=<<B><I>IC<SUB>t</SUB></I></B>>]
                  I2 [label=<<B><I>IC<SUB>t+1</SUB></I></B>>]
                  J1 [label=<<B><I>Emp<SUB>t</SUB></I></B>>]
                  J2 [label=<<B><I>Emp<SUB>t+1</SUB></I></B>>]
                  E1 [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E2 [label=<<B><I>FD<SUB>t+1</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y0 [label=<<B><I>MCS<SUB>b</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y2 [label=<<B><I>MCS<SUB>t+1</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E1 -> E2 
                  Y0 -> Y1
                  Y1 -> Y2
                  A1 -> A2
                  C1 -> C2
                  J1 -> J2
                  I1 -> I2

                  TI -> E1, E2, C1, C2, Y0, Y1, Y2, I1, I2, J1, J2

                  A1 -> C1, E1, Y1, J1, I1
                  A2 -> C2, E2, Y2, J2, I2

                  C1 -> E1, Y1, J1
                  C2 -> E2, Y2, J2
                  
                  I1 -> E1, Y1, J2
                  I2 -> E2, Y2

                  J1 -> E1, Y1, I1
                  J2 -> E2, Y2, I2

                  E1 -> C2
                  E1 -> Y1 
                  E2 -> Y2
                  
                  Y0 -> C1, J1, I1
                  Y1 -> C2, J2, I2
                  Y1 -> E2

                  {rank=min; TI}
                  {rank=same; A1, A2}
                  {rank=same; C1, C2}
                  {rank=same; E1, E2}
                  {rank=same; I1, I2}
                  {rank=same; J1, J2}
                  {rank=max; Y0, Y1, Y2}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point. The report retrieves cached information from `_targets.R` at the repo root.

```{r data_prep}
#| output: false
wide_data <- tar_read(wide_data, store = tar_store)
seed_random <- 20260522
```

# Multiple imputation (wide format)
```{r mice-gformula}
#| output: false
wide_mids <- tar_read(wide_mids, store = tar_store)
```

# `mice` positivity check
```{r mice-positivity}
#| output: false
#| include: true
#| eval: false

positivity_results <- function(mids_obj) {
  mice::complete(mids_obj, action = "long", include = FALSE) |>
    group_by(.imp,
             econ_dist_bin_0,
             econ_dist_bin_1,
             econ_dist_bin_2) |>
    summarise(n = n(), .groups = "drop") |>
    group_by(econ_dist_bin_0,
             econ_dist_bin_1,
             econ_dist_bin_2) |>
    summarise(
      mean_n = round(mean(n), 0),
      min_n  = min(n),
      max_n  = max(n),
      n_imp_with_zero = sum(n == 0),
      .groups = "drop"
    )
}

print(positivity_results(wide_mids))
```

# Main analysis: LTMLE with multiple imputation
The LTMLE sweep over the 8 regime × imputation pairs is computed by the targets DAG (`R/fit_ltmle_one.R` + `R/pool_ltmle.R`); each branch dispatches as a separate SLURM job via `future.batchtools` (template: `slurm.tmpl`). The pooled tibble is read in below.

```{r ltmle-mi}
#| output: false
#| include: true
ltmle_results <- tar_read(ltmle_results, store = tar_store)
```


# Sensitivity analyses: g-formula and IPTW

## G-formula via multiple imputation

### Imputing counterfactual outcomes under each regime using `gFormulaMI`
```{r gformula-mi}
#| include: false
#| output: false
mi_results <- tar_read(mi_results, store = tar_store)
```

## IPTW with multiple imputation

```{r iptw-mi}
#| output: false
#| include: false
iptw_fit  <- tar_read(iptw_fit, store = tar_store)
iptw_pool <- mice::pool(iptw_fit)
summary(iptw_pool)
```

### Table with results from all models
```{r table-results}
#| output: true
#| include: true
iptw_results <- tar_read(iptw_results, store = tar_store)
comparison   <- tar_read(comparison, store = tar_store)

comparison |> knitr::kable(digits = 3)
```

```{r plot_results}
#| output: true
#| include: true

# Combine results into a single data frame for plotting
results_comparison <- comparison |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("ltmle", "mi", "iptw_sw"),
                    labels = c("LTMLE", "G-Formula MI", "IPTW")),
    intervention = factor(intervention,
                          levels = c("0-0-0", "0-0-1", "0-1-0", "0-1-1", "1-0-0", "1-0-1", "1-1-0", "1-1-1"),
  )) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Intervention",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(39, 51, by = 2)) +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

print(results_comparison)

ggsave(
  filename = here::here("figs", "results_comparison.png"),
  plot     = results_comparison,
  width = 6, height = 8, dpi = 300
)

```