---
title: "Single point-treatment and outcome analysis"
subtitle: "Including only wave 3"
author: "Darwin Del Castillo"
format: 
  html:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from wave 3, which was collected from 2011 to 2012.

```{r setup}
#| output: false
here::i_am("07_single_treatment.qmd")

pacman::p_load(DiagrammeR,
               tidyverse,
               mice,
               targets,
               knitr)

# The targets store lives at the project root. Pass its absolute path to every
# tar_read() below so resolution doesn't depend on the working directory.
tar_store <- here::here("_targets")

source(here::here("R", "import_cleaning.R"))
source(here::here("R", "helpers.R"))
```

# Model specification: Two waves of data, time-invariant covariates + PCS
In this model, we include 8 waves of data (waves 3 to 10) and adjust for time-invariant covariates and time-varying confounders. The model is specified as follows: \
**Notation** \
- TI~cov~: Time-invariant covariates: \
    - Sex (`sex_dv`) \
    - Ethnicity (`race_dv`) \
    - Highest educational qualification (`hiqual_dv`) \
    - Region of residence (GOR) (`gor_dv`) \
    - Age at baseline (`age_dv_base`) \
- PCS~t-1~, PCS~t~, PCS~t+1~: Physical component summary score at time points *t-1*, *t*, and *t+1* \
- IC~t~, IC~t+1~, IC~t+2~: Intermediate confounders (lagged) at time points *t*, *t+1*, and *t+2*, which include: \
   - Number of dependents (`dnc`, 0, 1, 2+) \
   - Economic benefits receipt (`econ_benefits`, yes/no) \
   - Home ownership status (`home_owner`, yes/no) \
   - Marital status (`mastat_dv`, partnered vs not partnered) \

```{r model1}
#| echo: false
#| include: true
#| fig-cap: "Direct Acyclic Graph showing causal assumptions for the present analysis"
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=LR]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  C1 [label=<<B><I>CF<SUB>t-1</SUB></I></B>>]
                  C2 [label=<<B><I>CF<SUB>t</SUB></I></B>>]
                  I1 [label=<<B><I>IC<SUB>t</SUB></I></B>>]
                  I2 [label=<<B><I>IC<SUB>t+1</SUB></I></B>>]
                  J1 [label=<<B><I>Emp<SUB>t</SUB></I></B>>]
                  J2 [label=<<B><I>Emp<SUB>t+1</SUB></I></B>>]
                  E1 [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y0 [label=<<B><I>MCS<SUB>b</SUB></I></B>>]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  Y0 -> Y1
                  C1 -> C2
                  J1 -> J2
                  I1 -> I2

                  TI -> E1, C1, C2, Y0, Y1, I1, I2, J1, J2

                  C1 -> E1, Y1, J1
                  C2 -> J2
                  
                  I1 -> E1, Y1, J2

                  J1 -> E1, Y1, I1
                  J2 -> I2

                  E1 -> C2
                  E1 -> Y1 
                  
                  Y0 -> C1, J1, I1, E1
                  Y1 -> C2, J2, I2

                  {rank=min; E1}
                  {rank=same; TI}
                  {rank=same; C1, C2}
                  {rank=same; E1}
                  {rank=same; I1, I2}
                  {rank=same; J1, J2}
                  {rank=max; Y0, Y1}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point. The report retrieves cached information from `_targets.R` at the repo root.

```{r data_prep}
#| output: false
wide_data <- tar_read(wide_data, store = tar_store)
seed_random <- 42
```

# Multiple imputation (wide format)
```{r mice-gformula}
#| output: false
wide_mids <- tar_read(wide_mids, store = tar_store)
```

# Fit TMLE
We fit a single-point TMLE: the exposure is financial distress (`econ_dist_bin_0`) and the outcome is the SF-12 mental component summary (`sf12mcs_dv_0`). The table reports, pooled across imputations with Rubin's rules and in MCS points, the two counterfactual marginal means — mean SF-12 MCS if every person-wave were, versus were not, in financial distress — and their contrast (the ATE).

```{r fit-tmle}
tmle_results <- tar_read(tmle_results, store = tar_store)
knitr::kable(
  tmle_results,
  digits  = 3,
  caption = "Pooled single-point TMLE estimates: counterfactual marginal means (Distress / No distress) and their contrast (ATE), in SF-12 MCS points."
)
```

## Marginal means
Populational mean across the two counterfactuals.

```{r plot-tmle-means}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Single-point TMLE: pooled counterfactual marginal means of SF-12 MCS under financial distress vs no distress (95% CI)."
#| fig-width: 6
#| fig-height: 2.5

means_plot <- tmle_results |>
  filter(type == "mean") |>
  mutate(estimand = factor(estimand, levels = c("No distress", "Distress"))) |>
  ggplot(aes(x = effect, y = estimand, xmin = ll, xmax = ul, colour = estimand)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.2) +
  scale_colour_manual(values = c("No distress" = "grey40", "Distress" = "deepskyblue4"),
                      guide = "none") +
  labs(x = "Counterfactual mean SF-12 MCS (95% CI)", y = NULL) +
  theme(panel.grid.minor = element_blank())

print(means_plot)

ggsave(
  filename = here::here("figs", "tmle_means.png"),
  plot     = means_plot,
  width = 6, height = 2.5, dpi = 300
)
```

