---
title: "Single point-treatment and outcome analysis"
subtitle: "Including only wave 3"
author: "Darwin Del Castillo"
format: 
  html:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from wave 3, which was collected from 2011 to 2012.

```{r setup}
#| output: false
here::i_am("07_single_treatment.qmd")

pacman::p_load(DiagrammeR,
               tidyverse,
               mice,
               targets,
               knitr)

# The targets store lives at the project root. Pass its absolute path to every
# tar_read() below so resolution doesn't depend on the working directory.
tar_store <- here::here("_targets")

source(here::here("R", "import_cleaning.R"))
source(here::here("R", "helpers.R"))
```

# Model specification: One wave of data, time-invariant covariates + lagged confounders + baseline outcome + other confounders
In this model, we include 1 wave of data (wave 3) and adjust for time-invariant covariates and lagged confounders. The model is specified as follows:

**Notation**

- TI~cov~: Time-invariant covariates:
    - Sex (`sex_dv`)
    - Ethnicity (`race_dv`)
    - Highest educational qualification (`hiqual_dv`)
- C~l~: lagged confounders from the previous round
    - Number of dependents (`dnc`, 0, 1, 2+)
    - Economic benefits receipt (`econ_benefits`, yes/no)
    - Home ownership status (`home_owner`, yes/no)
    - Marital status (`mastat_dv`, partnered vs not partnered)
- C~f~: Other confounders (not lagged)
    - Region of residence (GOR) (`gor_dv`)
    - Age (`age_dv`)
- Job: Employment status (`econ_emp_bin`, employed vs not employed)
- Inc: Income (`log_income`)
- Y~0~: SF-12 mental component summary (MCS) at baseline (`sf12mcs_dv_base`)


```{r model1}
#| echo: false
#| include: true
#| fig-cap: "Direct Acyclic Graph showing causal assumptions for the present analysis"
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=LR]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  Cl [label=<<B><I>Con<SUB>l</SUB></I></B>>]
                  J [label=<<B><I>Job</I></B>>]
                  I [label=<<B><I>Inc</I></B>>]
                  E [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y0 [label=<<B><I>MCS<SUB>b</SUB></I></B>>]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E -> Y1
                  TI -> E, Y1, Cl, Y0
                  Cl -> E, Y1, J
                  J -> I, E
                  I -> E, Y1
                  Y0 -> Y1, Cl, E

                  {rank=min; E, TI}
                  {rank=same; Y0}
                  {rank=same; Cl}
                  {rank=same; J, I}
                  {rank=max; Y1}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point. The report retrieves cached information from `_targets.R` at the repo root.

```{r data_prep}
#| output: false
wide_data <- tar_read(wide_data, store = tar_store)
seed_random <- 42
```

# Multiple imputation (wide format)
```{r mice-gformula}
#| output: false
wide_mids <- tar_read(wide_mids, store = tar_store)
```

# Fit TMLE
We fit a single-point TMLE: the exposure is financial distress (`econ_dist_bin_0`) and the outcome is the SF-12 mental component summary (`sf12mcs_dv_0`). The table reports, pooled across imputations with Rubin's rules and in MCS points, the two counterfactual marginal means — mean SF-12 MCS if every person-wave were, versus were not, in financial distress — and their contrast (the ATE).

```{r fit-tmle}
tmle_results <- tar_read(tmle_results, store = tar_store)
knitr::kable(
  tmle_results,
  digits  = 3,
  caption = "Pooled single-point TMLE estimates: counterfactual marginal means (Distress / No distress) and their contrast (ATE), in SF-12 MCS points."
)
```

# Sensitivity analysis: g-formula with multiple imputation
We fit a g-formula model with multiple imputation, using the same model specification as the TMLE above. The following tables present the comparison between marginal means and ATE of the two counterfactuals estimated with TMLE and g-formula MI.

## Marginal means
```{r fit-sensitivity}
comparison <- tar_read(comparison, store = tar_store)

marginals <- comparison$marginal

knitr::kable(
  marginals,
  digits  = 3,
  caption = "Comparison of pooled single-point TMLE and g-formula MI estimates: counterfactual marginal means (Distress / No distress) and their contrast (ATE), in SF-12 MCS points."
)
```

## Average treatment effect (ATE)
```{r fit-sensitivity-ate}
ate <- comparison$ate

knitr::kable(
  ate,
  digits  = 3,
  caption = "Comparison of pooled single-point TMLE and g-formula MI estimates: average treatment effect (ATE) of financial distress on SF-12 MCS, in points."
)
```

# Graphs
## Marginal means
Populational mean across the two counterfactuals.
```{r plot-tmle-means}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Single-point TMLE: pooled counterfactual marginal means of SF-12 MCS under financial distress vs no distress (95% CI)."
#| fig-width: 6
#| fig-height: 2.5

results_comparison_marginal <- marginals |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("tmle", "mi"),
                    labels = c("TMLE", "G-Formula MI")),
    intervention = factor(intervention,
                          levels = c("No distress", "Distress"),
  )) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Intervention",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(39, 51, by = 2)) +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

print(results_comparison_marginal)

ggsave(
  filename = here::here("figs", "comparison_means.png"),
  plot     = results_comparison_marginal,
  width = 6, height = 2.5, dpi = 300
)
```

## Average treatment effect (ATE)
```{r plot-tmle-ate}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Single-point TMLE: pooled average treatment effect of financial distress on SF-12 MCS (95% CI)."
#| fig-width: 6
#| fig-height: 2.5

results_comparison_ate <- ate |>
  pivot_longer(
    -estimand,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("tmle", "mi"),
                    labels = c("TMLE", "G-Formula MI")),
    estimand = factor(estimand,
                      levels = c("Distress"),
  )) |>
  ggplot(aes(x = effect, y = estimand,
             colour = estimand,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Estimand",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(-4, -2, by = 0.5)) +
  labs(x = "Causal mean difference (95% CI)
  (Reference: No distress)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

print(results_comparison_ate)

ggsave(
  filename = here::here("figs", "comparison_ate.png"),
  plot     = results_comparison_ate,
  width = 6, height = 2.5, dpi = 300
)
```

# Secondary analysis: I-MAIHDA via TMLE

The intersectional strata (sex × education × ethnicity) are treated as clusters and the estimand becomes the **cluster-level PATE**, $E[Y_c(1) - Y_c(0)]$ where $Y_{c,j}$ is the mean SF-12 MCS in stratum $j$. Each stratum contributes equally regardless of how many people are in it, whereas the primary TMLE above weights each *person* equally. The two therefore only coincide when stratum size is uninformative about the effect — the divergence is the quantity of interest.

Inference treats the stratum, not the person, as the independent unit, so the effective sample size for the interval is the number of strata.

## Table
```{r fit-maihda}
#| output: true
maihda_results <- tar_read(maihda_results, store = tar_store)

knitr::kable(
  maihda_results,
  digits  = 3,
  caption = "Hierarchical (cluster) TMLE, pooled across imputations with Rubin's rules, in SF-12 MCS points. The estimator returns an influence curve for the contrast only, so the counterfactual means are reported as point estimates without intervals. `df` is the Barnard-Rubin degrees of freedom with the number of strata as the complete-data sample size."
)
```

## Graph
```{r plot-maihda}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Financial distress and SF-12 MCS, estimated two ways: the primary single-point TMLE (each person weighted equally) and the hierarchical cluster TMLE with intersectional strata as clusters (each stratum weighted equally). Intervals are 95%; the cluster-TMLE interval uses a t cutoff on the number of strata, so it is wider by construction."
#| fig-width: 7
#| fig-height: 4

maihda_plot_data <- bind_rows(
  tmle_results |>
    select(estimand, type, effect, ll, ul) |>
    mutate(method = "TMLE (person-weighted)"),
  maihda_results |>
    select(estimand, type, effect, ll, ul) |>
    mutate(method = "Cluster TMLE (stratum-weighted)")
) |>
  mutate(
    method   = factor(method, levels = c("TMLE (person-weighted)",
                                         "Cluster TMLE (stratum-weighted)")),
    estimand = factor(estimand, levels = c("No distress", "Distress", "ATE")),
    # Means (~50 MCS points) and the contrast (~-2) cannot share an x axis.
    panel    = factor(ifelse(type == "ate", "Causal contrast (ATE)",
                             "Counterfactual marginal means"),
                      levels = c("Counterfactual marginal means",
                                 "Causal contrast (ATE)"))
  )

results_maihda <- maihda_plot_data |>
  ggplot(aes(x = effect, y = estimand, colour = method,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_vline(data = data.frame(panel = factor("Causal contrast (ATE)",
                                              levels = levels(maihda_plot_data$panel))),
             aes(xintercept = 0), linetype = "dashed", colour = "grey50") +
  geom_point(size = 2.5, position = position_dodge(width = 0.5)) +
  geom_errorbar(orientation = "y", width = 0.3, na.rm = TRUE,
                position = position_dodge(width = 0.5)) +
  scale_colour_brewer(palette = "Set1", name = "Estimator", guide = "legend") +
  # "free" (not "free_x"): the means and the contrast need separate x ranges, and free y
  # also drops the estimand levels that don't belong to each panel.
  facet_wrap(~panel, ncol = 1, scales = "free") +
  labs(x = "SF-12 MCS points (95% CI)", y = NULL) +
  theme(panel.grid.minor = element_blank(),
        legend.position  = "bottom")

print(results_maihda)

ggsave(
  filename = here::here("figs", "maihda_comparison.png"),
  plot     = results_maihda,
  width = 7, height = 4, dpi = 300
)
```


