---
title: "Single point-treatment and outcome analysis"
subtitle: "Including only wave 3"
author: "Darwin Del Castillo"
format: 
  html:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from wave 3, which was collected from 2011 to 2012.

```{r setup}
#| output: false
here::i_am("07_single_treatment.qmd")

pacman::p_load(DiagrammeR,
               tidyverse,
               mice,
               targets,
               knitr)

# The targets store lives at the project root. Pass its absolute path to every
# tar_read() below so resolution doesn't depend on the working directory.
tar_store <- here::here("_targets")

source(here::here("R", "import_cleaning.R"))
source(here::here("R", "helpers.R"))
```

# Model specification: One wave of data, time-invariant covariates + lagged confounders + baseline outcome + other confounders
In this model, we include 1 wave of data (wave 3) and adjust for time-invariant covariates and lagged confounders. The model is specified as follows:

**Notation**

- TI~cov~: Time-invariant covariates:
    - Sex (`sex_dv`)
    - Ethnicity (`race_dv`)
    - Highest educational qualification (`hiqual_dv`)
- C~l~: lagged confounders from the previous round
    - Number of dependents (`dnc`, 0, 1, 2+)
    - Economic benefits receipt (`econ_benefits`, yes/no)
    - Home ownership status (`home_owner`, yes/no)
    - Marital status (`mastat_dv`, partnered vs not partnered)
- C~f~: Other confounders (not lagged)
    - Region of residence (GOR) (`gor_dv`)
    - Age (`age_dv`)
- Job: Employment status (`econ_emp_bin`, employed vs not employed)
- Inc: Income (`log_income`)
- Y~0~: SF-12 mental component summary (MCS) at baseline (`sf12mcs_dv_base`)


```{r model1}
#| echo: false
#| include: true
#| fig-cap: "Direct Acyclic Graph showing causal assumptions for the present analysis"
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=LR]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  Cl [label=<<B><I>Con<SUB>l</SUB></I></B>>]
                  J [label=<<B><I>Job</I></B>>]
                  I [label=<<B><I>Inc</I></B>>]
                  E [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y0 [label=<<B><I>MCS<SUB>b</SUB></I></B>>]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E -> Y1
                  TI -> E, Y1, Cl, Y0
                  Cl -> E, Y1, J
                  J -> I, E
                  I -> E, Y1
                  Y0 -> Y1, Cl, E

                  {rank=min; E, TI}
                  {rank=same; Y0}
                  {rank=same; Cl}
                  {rank=same; J, I}
                  {rank=max; Y1}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point. The report retrieves cached information from `_targets.R` at the repo root.

```{r data_prep}
#| output: false
wide_data <- tar_read(wide_data, store = tar_store)
seed_random <- 42
```

# Multiple imputation (wide format)
```{r mice-gformula}
#| output: false
wide_mids <- tar_read(wide_mids, store = tar_store)
```

# Fit TMLE
We fit a single-point TMLE: the exposure is financial distress (`econ_dist_bin_0`) and the outcome is the SF-12 mental component summary (`sf12mcs_dv_0`). The table reports, pooled across imputations with Rubin's rules and in MCS points, the two counterfactual marginal means — mean SF-12 MCS if every person-wave were, versus were not, in financial distress — and their contrast (the ATE).

```{r fit-tmle}
tmle_results <- tar_read(tmle_results, store = tar_store)
knitr::kable(
  tmle_results,
  digits  = 3,
  caption = "Pooled single-point TMLE estimates: counterfactual marginal means (Distress / No distress) and their contrast (ATE), in SF-12 MCS points."
)
```

# Sensitivity analysis: g-formula with multiple imputation
We fit a g-formula model with multiple imputation, using the same model specification as the TMLE above. The following tables present the comparison between marginal means and ATE of the two counterfactuals estimated with TMLE and g-formula MI.

## Marginal means
```{r fit-sensitivity}
comparison <- tar_read(comparison, store = tar_store)

marginals <- comparison$marginal

knitr::kable(
  marginals,
  digits  = 3,
  caption = "Comparison of pooled single-point TMLE and g-formula MI estimates: counterfactual marginal means (Distress / No distress) and their contrast (ATE), in SF-12 MCS points."
)
```

## Average treatment effect (ATE)
```{r fit-sensitivity-ate}
ate <- comparison$ate

knitr::kable(
  ate,
  digits  = 3,
  caption = "Comparison of pooled single-point TMLE and g-formula MI estimates: average treatment effect (ATE) of financial distress on SF-12 MCS, in points."
)
```

# Graphs
## Marginal means
Populational mean across the two counterfactuals.
```{r plot-tmle-means}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Single-point TMLE: pooled counterfactual marginal means of SF-12 MCS under financial distress vs no distress (95% CI)."
#| fig-width: 6
#| fig-height: 2.5

results_comparison_marginal <- marginals |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("tmle", "mi"),
                    labels = c("TMLE", "G-Formula MI")),
    intervention = factor(intervention,
                          levels = c("No distress", "Distress"),
  )) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Intervention",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(39, 51, by = 2)) +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

print(results_comparison_marginal)

ggsave(
  filename = here::here("figs", "comparison_means.png"),
  plot     = results_comparison_marginal,
  width = 6, height = 2.5, dpi = 300
)
```

## Average treatment effect (ATE)
```{r plot-tmle-ate}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Single-point TMLE: pooled average treatment effect of financial distress on SF-12 MCS (95% CI)."
#| fig-width: 6
#| fig-height: 2.5

results_comparison_ate <- ate |>
  pivot_longer(
    -estimand,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("tmle", "mi"),
                    labels = c("TMLE", "G-Formula MI")),
    estimand = factor(estimand,
                      levels = c("Distress"),
  )) |>
  ggplot(aes(x = effect, y = estimand,
             colour = estimand,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Estimand",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(-4, -2, by = 0.5)) +
  labs(x = "Causal mean difference (95% CI)
  (Reference: No distress)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

print(results_comparison_ate)

ggsave(
  filename = here::here("figs", "comparison_ate.png"),
  plot     = results_comparison_ate,
  width = 6, height = 2.5, dpi = 300
)
```

# Secondary analysis: DR-learner GATEs by intersectional stratum

The marginal TMLE above gives one average effect. This section asks *for whom* the effect differs, estimating a **group average treatment effect** for each of the 12 intersectional strata $j$ (sex × ethnicity × education):

$$\text{GATE}_j = E[\,Y(0) - Y(1) \mid S = j\,]$$

**Sign convention.** $Y(0)$ is the no-distress potential outcome and financial distress lowers MCS, so $\text{GATE}_j > 0$ reads as **SF-12 MCS points *lost* to financial distress** in stratum $j$. Because the three stratum variables are already in the adjustment set $W$, no new identification assumption is introduced. Each person's doubly-robust (AIPW) pseudo-outcome is built from independently cross-fit SuperLearner nuisances and averaged within stratum via a saturated projection with HC2 standard errors; results are pooled across imputations with Rubin's rules.

## Effects by stratum

```{r cate-table}
#| output: true
cate_results <- tar_read(cate_results, store = tar_store)

cate_results$gate |>
  arrange(desc(estimate)) |>
  select(strata_label, n_j_min, n_j_max, pct_exposed, min_g, median_g,
         share_g_at_bound, estimate, ll, ul) |>
  knitr::kable(
    digits    = c(NA, 0, 0, 1, 3, 3, 3, 2, 2, 2),
    col.names = c("Stratum (sex:race:education)", "n (min)", "n (max)",
                  "% exposed", "min ĝ", "median ĝ", "share ĝ at bound",
                  "GATE", "LL", "UL"),
    caption   = "DR-learner group average treatment effects by intersectional stratum, pooled across imputations with Rubin's rules. GATE is E[Y(0) − Y(1) | S = j]: positive values are SF-12 MCS points LOST to financial distress. n (min/max) is the stratum-size range across imputations; the ĝ diagnostics flag cells whose estimate leans on extrapolated outcome predictions rather than overlap."
  )
```

## Is the heterogeneity real?

Three complementary answers: the Chernozhukov *best linear predictor* (BLP) coefficient $\beta_2$ (on the conventional $Y(1)-Y(0)$ scale; $\beta_2 \neq 0$ means the cell-level variation is signal rather than noise), the omnibus Wald test of $\text{GATE}_1 = \dots = \text{GATE}_{12}$, and — as a specification check — the DR score's marginal mean against the sign-flipped TMLE ATE.

```{r cate-heterogeneity}
#| output: true
cate_results$blp |>
  knitr::kable(
    digits    = 3,
    col.names = c("Term", "Estimate", "SE", "LL", "UL"),
    caption = "BLP heterogeneity test, pooled across imputations (HC3 SEs within imputation). beta1_ate is the marginal effect on the conventional Y(1) − Y(0) scale and should sit near the TMLE ATE in magnitude; it is therefore opposite in sign to the DR-score marginal mean (cate_results$ate$estimate, shown in the sanity check below), which is on the Y(0) − Y(1) scale. beta2_hte ≠ 0 is evidence that the 12 cells genuinely differ."
  )

# df1 (numerator df = number of strata - 1) is expected constant across
# imputations; pool_cate.R's own header flags a differing stratum count as
# theoretically possible, so assert rather than silently taking the first.
stopifnot(
  "wald test df1 differs across imputations -- stratum count is not stable" =
    dplyr::n_distinct(cate_results$wald$df1) == 1
)

cate_results$wald |>
  summarise(
    `median F` = median(F),
    df1        = first(df1),
    `median p` = median(p),
    `min p`    = min(p),
    `max p`    = max(p)
  ) |>
  knitr::kable(
    digits  = 3,
    caption = "Omnibus Wald test of GATE_1 = ... = GATE_12. F statistics are not Rubin-pooled; the median and range across imputations are shown."
  )
```

```{r cate-sanity}
#| output: true
tmle_ate_row <- tmle_results |> filter(type == "ate")

tibble(
  quantity = c("Pooled mean(ψ) — DR score, Y(0) − Y(1) scale",
               "−1 × TMLE ATE (tmle_results)"),
  estimate = c(cate_results$ate$estimate, -tmle_ate_row$effect),
  ll       = c(cate_results$ate$ll, -tmle_ate_row$ul),
  ul       = c(cate_results$ate$ul, -tmle_ate_row$ll)
) |>
  knitr::kable(
    digits  = 3,
    caption = "Sanity row: the DR score's marginal mean against the sign-flipped TMLE ATE. The two estimators share an estimand but no nuisance fits, so close agreement is expected and material disagreement indicates a nuisance-specification bug."
  )
```

## Caterpillar plot

```{r plot-cate-caterpillar}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Effect of financial distress on SF-12 MCS by intersectional stratum, E[Y(0) − Y(1) | S] with 95% CI: positive values are MCS points lost to distress. Dashed line: the marginal TMLE ATE, sign-flipped to match. Dotted line: zero."
#| fig-width: 7
#| fig-height: 4.5

ref_ate <- -(tmle_results |> filter(type == "ate") |> pull(effect))

cate_caterpillar <- cate_results$gate |>
  mutate(strata_label = forcats::fct_reorder(factor(strata_label), estimate)) |>
  ggplot(aes(x = estimate, y = strata_label)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = "dotted", colour = "grey60") +
  geom_vline(xintercept = ref_ate, linetype = "dashed", colour = "grey30") +
  geom_pointrange(aes(xmin = ll, xmax = ul), size = 0.4) +
  labs(x = "SF-12 MCS points lost to financial distress (95% CI)",
       y = NULL) +
  theme(panel.grid.minor = element_blank())

print(cate_caterpillar)

ggsave(
  filename = here::here("figs", "cate_caterpillar.png"),
  plot     = cate_caterpillar,
  width = 7, height = 4.5, dpi = 300
)
```

```{r cate-thinnest}
#| output: false
# Computed, not hardcoded: Workstream B re-runs the imputation, so a fixed
# prose number for "the thinnest cell" would silently go stale. Derived from
# the 12-row gate tibble so it tracks whatever data actually lands.
thin <- cate_results$gate |> dplyr::slice_min(n_j_min, n = 1, with_ties = FALSE)
```

Three cautions bound this section. First, congeniality between the imputation model and a between-stratum contrast analysis is improved by the saturated $A \times S$ imputation terms but not exact — the analysis nuisances are nonparametric in all of $W$. Second, the thinnest cell (n ≈ `r thin$n_j_min` at ~`r round(thin$pct_exposed)`% exposure) rests on roughly `r round(thin$n_j_min * thin$pct_exposed / 100)` exposed observations; its wide interval and its ĝ diagnostics keep that visible. Third, the 12 projection SEs ignore cross-cell correlation induced by the shared nuisance fits, as is conventional.

