---
title: "Single point-treatment and outcome analysis"
subtitle: "Including only wave 3"
author: "Darwin Del Castillo"
format: 
  html:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from wave 3, which was collected from 2011 to 2012.

```{r setup}
#| output: false
here::i_am("07_single_treatment.qmd")

pacman::p_load(DiagrammeR,
               tidyverse,
               mice,
               targets,
               knitr,
               ggrepel)

# The targets store lives at the project root. Pass its absolute path to every
# tar_read() below so resolution doesn't depend on the working directory.
tar_store <- here::here("_targets")

source(here::here("R", "import_cleaning.R"))
source(here::here("R", "helpers.R"))
```

# Model specification: One wave of data, time-invariant covariates + lagged confounders + baseline outcome + other confounders
In this model, we include 1 wave of data (wave 3) and adjust for time-invariant covariates and lagged confounders. The model is specified as follows:

**Notation**

- TI~cov~: Time-invariant covariates:
    - Sex (`sex_dv`)
    - Ethnicity (`race_dv`)
    - Highest educational qualification (`hiqual_dv`)
- C~l~: lagged confounders from the previous round
    - Number of dependents (`dnc`, 0, 1, 2+)
    - Economic benefits receipt (`econ_benefits`, yes/no)
    - Home ownership status (`home_owner`, yes/no)
    - Marital status (`mastat_dv`, partnered vs not partnered)
- C~f~: Other confounders (not lagged)
    - Region of residence (GOR) (`gor_dv`)
    - Age (`age_dv`)
- Job: Employment status (`econ_emp_bin`, employed vs not employed)
- Inc: Income (`log_income`)
- Y~0~: SF-12 mental component summary (MCS) at baseline (`sf12mcs_dv_base`)


```{r model1}
#| echo: false
#| include: true
#| fig-cap: "Direct Acyclic Graph showing causal assumptions for the present analysis"
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=LR]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  Cl [label=<<B><I>Con<SUB>l</SUB></I></B>>]
                  J [label=<<B><I>Job</I></B>>]
                  I [label=<<B><I>Inc</I></B>>]
                  E [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y0 [label=<<B><I>MCS<SUB>b</SUB></I></B>>]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E -> Y1
                  TI -> E, Y1, Cl, Y0
                  Cl -> E, Y1, J
                  J -> I, E
                  I -> E, Y1
                  Y0 -> Y1, Cl, E

                  {rank=min; E, TI}
                  {rank=same; Y0}
                  {rank=same; Cl}
                  {rank=same; J, I}
                  {rank=max; Y1}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point. The report retrieves cached information from `_targets.R` at the repo root.

```{r data_prep}
#| output: false
wide_data <- tar_read(wide_data, store = tar_store)
seed_random <- 42
```

# Multiple imputation (wide format)
```{r mice-gformula}
#| output: false
wide_mids <- tar_read(wide_mids, store = tar_store)
```

# Fit TMLE
We fit a single-point TMLE: the exposure is financial distress (`econ_dist_bin_0`) and the outcome is the SF-12 mental component summary (`sf12mcs_dv_0`). The table reports, pooled across imputations with Rubin's rules and in MCS points, the two counterfactual marginal means — mean SF-12 MCS if every person-wave were, versus were not, in financial distress — and their contrast (the ATE).

```{r fit-tmle}
tmle_results <- tar_read(tmle_results, store = tar_store)
knitr::kable(
  tmle_results,
  digits  = 3,
  caption = "Pooled single-point TMLE estimates: counterfactual marginal means (Distress / No distress) and their contrast (ATE), in SF-12 MCS points."
)
```

# Sensitivity analysis: g-formula with multiple imputation
We fit a g-formula model with multiple imputation, using the same model specification as the TMLE above. The following tables present the comparison between marginal means and ATE of the two counterfactuals estimated with TMLE and g-formula MI.

## Marginal means
```{r fit-sensitivity}
comparison <- tar_read(comparison, store = tar_store)

marginals <- comparison$marginal

knitr::kable(
  marginals,
  digits  = 3,
  caption = "Comparison of pooled single-point TMLE and g-formula MI estimates: counterfactual marginal means (Distress / No distress) and their contrast (ATE), in SF-12 MCS points."
)
```

## Average treatment effect (ATE)
```{r fit-sensitivity-ate}
ate <- comparison$ate

knitr::kable(
  ate,
  digits  = 3,
  caption = "Comparison of pooled single-point TMLE and g-formula MI estimates: average treatment effect (ATE) of financial distress on SF-12 MCS, in points."
)
```

# Graphs
## Marginal means
Populational mean across the two counterfactuals.
```{r plot-tmle-means}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Single-point TMLE: pooled counterfactual marginal means of SF-12 MCS under financial distress vs no distress (95% CI)."
#| fig-width: 6
#| fig-height: 2.5

results_comparison_marginal <- marginals |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("tmle", "mi"),
                    labels = c("TMLE", "G-Formula MI")),
    intervention = factor(intervention,
                          levels = c("No distress", "Distress"),
  )) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Intervention",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(39, 51, by = 2)) +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

print(results_comparison_marginal)

ggsave(
  filename = here::here("figs", "comparison_means.png"),
  plot     = results_comparison_marginal,
  width = 6, height = 2.5, dpi = 300
)
```

## Average treatment effect (ATE)
```{r plot-tmle-ate}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Single-point TMLE: pooled average treatment effect of financial distress on SF-12 MCS (95% CI)."
#| fig-width: 6
#| fig-height: 2.5

results_comparison_ate <- ate |>
  pivot_longer(
    -estimand,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("tmle", "mi"),
                    labels = c("TMLE", "G-Formula MI")),
    estimand = factor(estimand,
                      levels = c("Distress"),
  )) |>
  ggplot(aes(x = effect, y = estimand,
             colour = estimand,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Estimand",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(-4, -2, by = 0.5)) +
  labs(x = "Causal mean difference (95% CI)
  (Reference: No distress)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())

print(results_comparison_ate)

ggsave(
  filename = here::here("figs", "comparison_ate.png"),
  plot     = results_comparison_ate,
  width = 6, height = 2.5, dpi = 300
)
```

# Secondary analysis: DR-learner GATEs by intersectional stratum

The marginal TMLE above gives one average effect. This section asks *for whom* the effect differs, estimating a **group average treatment effect** for each of the 12 intersectional strata $j$ (sex × ethnicity × education):

$$\text{GATE}_j = E[\,Y(0) - Y(1) \mid S = j\,]$$

**Sign convention.** $Y(0)$ is the no-distress potential outcome and financial distress lowers MCS, so $\text{GATE}_j > 0$ reads as **SF-12 MCS points *lost* to financial distress** in stratum $j$. Because the three stratum variables are already in the adjustment set $W$, no new identification assumption is introduced. Each person's doubly-robust (AIPW) pseudo-outcome is built from independently cross-fit SuperLearner nuisances and averaged within stratum via a saturated projection with HC2 standard errors; results are pooled across imputations with Rubin's rules.

## Effects by stratum

```{r cate-table}
#| output: true
cate_results <- tar_read(cate_results, store = tar_store)

cate_results$gate |>
  arrange(desc(estimate)) |>
  select(strata_label, n_j_min, n_j_max, pct_exposed, min_g, median_g,
         share_g_at_bound, estimate, ll, ul) |>
  knitr::kable(
    digits    = c(NA, 0, 0, 1, 3, 3, 3, 2, 2, 2),
    col.names = c("Stratum (sex:race:education)", "n (min)", "n (max)",
                  "% exposed", "min ĝ", "median ĝ", "share ĝ at bound",
                  "GATE", "LL", "UL"),
    caption   = "DR-learner group average treatment effects by intersectional stratum, pooled across imputations with Rubin's rules. GATE is E[Y(0) − Y(1) | S = j]: positive values are SF-12 MCS points LOST to financial distress. n (min/max) is the stratum-size range across imputations; the ĝ diagnostics flag cells whose estimate leans on extrapolated outcome predictions rather than overlap."
  )
```

## Is the heterogeneity real?

Three complementary answers: the Chernozhukov *best linear predictor* (BLP) coefficient $\beta_2$ (on the conventional $Y(1)-Y(0)$ scale; $\beta_2 \neq 0$ means the cell-level variation is signal rather than noise), the omnibus Wald test of $\text{GATE}_1 = \dots = \text{GATE}_{12}$, and — as a specification check — the DR score's marginal mean against the sign-flipped TMLE ATE.

```{r cate-heterogeneity}
#| output: true
cate_results$blp |>
  knitr::kable(
    digits    = 3,
    col.names = c("Term", "Estimate", "SE", "LL", "UL"),
    caption = "BLP heterogeneity test, pooled across imputations (HC3 SEs within imputation). beta1_ate is the marginal effect on the conventional Y(1) − Y(0) scale and should sit near the TMLE ATE in magnitude; it is therefore opposite in sign to the DR-score marginal mean (cate_results$ate$estimate, shown in the sanity check below), which is on the Y(0) − Y(1) scale. beta2_hte ≠ 0 is evidence that the 12 cells genuinely differ."
  )

# df1 (numerator df = number of strata - 1) is expected constant across
# imputations; pool_cate.R's own header flags a differing stratum count as
# theoretically possible, so assert rather than silently taking the first.
stopifnot(
  "wald test df1 differs across imputations -- stratum count is not stable" =
    dplyr::n_distinct(cate_results$wald$df1) == 1
)

cate_results$wald |>
  summarise(
    `median F` = median(F),
    df1        = first(df1),
    `median p` = median(p),
    `min p`    = min(p),
    `max p`    = max(p)
  ) |>
  knitr::kable(
    digits  = 3,
    caption = "Omnibus Wald test of GATE_1 = ... = GATE_12. F statistics are not Rubin-pooled; the median and range across imputations are shown."
  )
```

```{r cate-sanity}
#| output: true
tmle_ate_row <- tmle_results |> filter(type == "ate")

tibble(
  quantity = c("Pooled mean(ψ) — DR score, Y(0) − Y(1) scale",
               "−1 × TMLE ATE (tmle_results)"),
  estimate = c(cate_results$ate$estimate, -tmle_ate_row$effect),
  ll       = c(cate_results$ate$ll, -tmle_ate_row$ul),
  ul       = c(cate_results$ate$ul, -tmle_ate_row$ll)
) |>
  knitr::kable(
    digits  = 3,
    caption = "Sanity row: the DR score's marginal mean against the sign-flipped TMLE ATE. The two estimators share an estimand but no nuisance fits, so close agreement is expected and material disagreement indicates a nuisance-specification bug."
  )
```

## Caterpillar plot

```{r plot-cate-caterpillar}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Effect of financial distress on SF-12 MCS by intersectional stratum, E[Y(0) − Y(1) | S] with 95% CI: positive values are MCS points lost to distress. Dashed line: the marginal TMLE ATE, sign-flipped to match. Dotted line: zero."
#| fig-width: 7
#| fig-height: 4.5

ref_ate <- -(tmle_results |> filter(type == "ate") |> pull(effect))

cate_caterpillar <- cate_results$gate |>
  mutate(strata_label = forcats::fct_reorder(factor(strata_label), estimate)) |>
  ggplot(aes(x = estimate, y = strata_label)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = "dotted", colour = "grey60") +
  geom_vline(xintercept = ref_ate, linetype = "dashed", colour = "grey30") +
  geom_pointrange(aes(xmin = ll, xmax = ul), size = 0.4) +
  labs(x = "SF-12 MCS points lost to financial distress (95% CI)",
       y = NULL) +
  theme(panel.grid.minor = element_blank())

print(cate_caterpillar)

ggsave(
  filename = here::here("figs", "cate_caterpillar.png"),
  plot     = cate_caterpillar,
  width = 7, height = 4.5, dpi = 300
)
```

```{r cate-thinnest}
#| output: false
# Computed, not hardcoded: Workstream B re-runs the imputation, so a fixed
# prose number for "the thinnest cell" would silently go stale. Derived from
# the 12-row gate tibble so it tracks whatever data actually lands.
thin <- cate_results$gate |> dplyr::slice_min(n_j_min, n = 1, with_ties = FALSE)
```

Three cautions bound this section. First, congeniality between the imputation model and a between-stratum contrast analysis is improved by the saturated $A \times S$ imputation terms but not exact — the analysis nuisances are nonparametric in all of $W$. Second, the thinnest cell (n ≈ `r thin$n_j_min` at ~`r round(thin$pct_exposed)`% exposure) rests on roughly `r round(thin$n_j_min * thin$pct_exposed / 100)` exposed observations; its wide interval and its ĝ diagnostics keep that visible. Third, the 12 projection SEs ignore cross-cell correlation induced by the shared nuisance fits, as is conventional.

# Decomposition: additive main effects vs intersectional interaction

The previous section establishes *that* the 12 strata differ. This one asks *how much* of that
difference additive main effects can account for. The second stage treats the 12 GATEs as
`GATE_j = x_j'β + u_j + e_j` with `u_j ~ N(0, τ²)` and `e_j ~ N(0, SE_j²)`, fitting it twice: once with
no moderators and once with additive sex, race and education. `PCV = (τ²_null − τ²_main)/τ²_null` is the
share of between-stratum effect variance the additive surface explains; the remainder is intersectional.
All quantities keep the `Y(0) − Y(1)` convention — positive is MCS points lost to distress.

```{r maihda-read}
#| output: false
gate_meta <- tar_read(gate_meta_results, store = tar_store)
```

## Two routes to the same question

The DR-learner section already tested heterogeneity through the best linear predictor and an omnibus
Wald test on the saturated projection. The random-effects stage tests it again through a completely
separate mechanism — a variance component estimated by REML from the 12 estimates and their standard
errors. The two share no machinery beyond the GATEs themselves, so agreement between them is
corroboration rather than restatement.

```{r maihda-two-routes}
#| output: true
blp2 <- cate_results$blp |> dplyr::filter(term == "beta2_hte")

tibble::tibble(
  route = c("DR-learner", "DR-learner",
            "Random effects", "Random effects", "Random effects"),
  statistic = c("BLP beta2 (heterogeneity slope)",
                "Omnibus Wald F, median across imputations",
                "tau: SD of true stratum effects (MCS points)",
                "I^2 with no moderators (%)",
                "Q with no moderators, median across imputations"),
  estimate = c(blp2$estimate,
               stats::median(cate_results$wald$F),
               sqrt(max(0, gate_meta$scalars$tau2_null)),
               gate_meta$scalars$I2_null,
               stats::median(gate_meta$tests$QE_null)),
  ll = c(blp2$ll, NA, sqrt(max(0, gate_meta$scalars$tau2_null_ll)), NA, NA),
  ul = c(blp2$ul, NA, sqrt(max(0, gate_meta$scalars$tau2_null_ul)), NA, NA)
) |>
  knitr::kable(
    digits  = 3,
    caption = "Heterogeneity across the 12 intersectional strata, assessed two independent ways. The DR-learner statistics come from the saturated projection of the pseudo-outcome; the random-effects statistics come from a REML variance component fitted to the 12 (GATE, SE) pairs. They share no estimation machinery, so consistent verdicts are mutual corroboration. F and Q do not pool under Rubin's rules, hence medians."
  )
```

## Whose average? Equal-stratum weight vs person weight

```{r maihda-weighting}
#| output: true
tmle_ate <- tmle_results |> dplyr::filter(type == "ate")

tibble::tibble(
  quantity = c("Equal-stratum-weight mean effect (mu)",
               "Person-weighted DR-score marginal effect",
               "Person-weighted TMLE ATE (sign-flipped to match)"),
  estimate = c(gate_meta$scalars$mu, cate_results$ate$estimate, -tmle_ate$effect),
  ll       = c(gate_meta$scalars$mu_ll, cate_results$ate$ll, -tmle_ate$ul),
  ul       = c(gate_meta$scalars$mu_ul, cate_results$ate$ul, -tmle_ate$ll)
) |>
  knitr::kable(
    digits  = 3,
    caption = "The same effect under two weighting schemes, both on the Y(0) - Y(1) scale (MCS points lost to distress). mu weights every intersectional stratum equally; the marginal estimates weight by population share. A gap between them is not disagreement between estimators -- it says the effect differs across strata whose sizes differ, and here the smaller non-white strata carry smaller effects."
  )
```

## Decomposition

```{r maihda-decomposition}
#| output: true
tibble::tibble(
  quantity = c("tau^2, no moderators", "tau^2, additive main effects",
               "I^2, no moderators (%)", "I^2, additive (%)",
               "PCV (share explained by main effects)"),
  estimate = c(gate_meta$scalars$tau2_null, gate_meta$scalars$tau2_main,
               gate_meta$scalars$I2_null,   gate_meta$scalars$I2_main,
               gate_meta$scalars$pcv),
  ll = c(gate_meta$scalars$tau2_null_ll, gate_meta$scalars$tau2_main_ll,
         NA, NA, NA),
  ul = c(gate_meta$scalars$tau2_null_ul, gate_meta$scalars$tau2_main_ul,
         NA, NA, NA)
) |>
  knitr::kable(
    digits  = 3,
    caption = "Random-effects decomposition of the 12 intersectional GATEs, pooled across imputations (shrink-then-pool: the fits run within each imputation, Rubin's rules after). tau^2 is between-stratum variance of true effects in squared MCS points. PCV is reported without an interval: with 12 cells its sampling uncertainty is not credibly quantifiable, as is conventional in MAIHDA."
  )
```

```{r maihda-moderators}
#| output: true
# Reference categories are set by factor()'s default alphabetical ordering in
# .gate_factors() (R/fit_gate_meta.R): Female < Male, Non.white < White,
# High < Low < Medium -- so the reference cell is Female:Non.white:High.
term_order  <- c("(Intercept)", "sexMale", "raceWhite", "educLow", "educMedium")
term_labels <- c("Intercept (Female, Non.white, High)",
                 "Male (ref: Female)",
                 "White (ref: Non.white)",
                 "Low education (ref: High)",
                 "Medium education (ref: High)")

gate_meta$coefs |>
  dplyr::mutate(term = factor(term, levels = term_order, labels = term_labels)) |>
  dplyr::arrange(term) |>
  knitr::kable(
    digits    = 3,
    col.names = c("Term", "Estimate", "SE", "LL", "UL"),
    caption   = "Additive main-effect moderators of the stratum-specific effect, pooled. Reference categories: Female (sex), Non.white (race), High (education) -- so the intercept is the additive-model prediction for the Female:Non.white:High stratum. These are contrasts of GATEs, not of outcomes: a coefficient is how many more MCS points distress costs that group relative to the reference."
  )
```

```{r maihda-tests}
#| output: true
tibble::tibble(
  statistic = c("Moderator test F (median)", "Moderator test p (median)",
                "Residual QE, additive model (median)",
                "Imputations with PCV < 0"),
  value = c(stats::median(gate_meta$tests$QM),
            stats::median(gate_meta$tests$QMp),
            stats::median(gate_meta$tests$QE_main),
            sum(gate_meta$tests$pcv < 0, na.rm = TRUE))
) |>
  knitr::kable(
    digits  = 3,
    caption = "Per-imputation test statistics. F and Q statistics do not pool under Rubin's rules, so medians across imputations are reported. A non-significant moderator test alongside surviving residual heterogeneity is the signature of effect variation that is intersectional rather than additive."
  )
```

## Additive prediction vs estimated effect

```{r maihda-additivity}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Additive main-effects prediction against the estimated GATE for each intersectional stratum. Points on the 45-degree line are fully explained by additive sex, race and education; vertical departures are the intersectional component. Y(0) - Y(1) scale: positive is MCS points lost to distress."
#| fig-width: 6
#| fig-height: 5
ggplot(gate_meta$cells, aes(x = additive, y = gate)) +
  theme_bw() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = gate_ll, ymax = gate_ul), size = 0.3) +
  ggrepel::geom_text_repel(aes(label = strata_label), size = 2.6,
                           max.overlaps = 20, seed = 1) +
  labs(x = "Additive main-effects prediction (MCS points)",
       y = "Estimated GATE (MCS points)") +
  theme(panel.grid.minor = element_blank())
```

## Raw effects and shrunken estimates

```{r maihda-caterpillar}
#| output: true
#| include: true
#| eval: true
#| fig-cap: "Raw DR-learner GATE (open circles, 95% CI) and the shrunken random-effects estimate (filled) for each stratum. Dashed line: the equal-stratum-weight mean effect mu. Shrunken values are predictions under the assumption that stratum effects are exchangeable draws, and pull toward the additive surface by construction -- every stratum-specific claim in the text quotes the RAW estimate."
#| fig-width: 7
#| fig-height: 4.5
ord <- gate_meta$cells |> dplyr::arrange(gate) |> dplyr::pull(strata_label)

gate_meta$cells |>
  dplyr::mutate(strata_label = factor(strata_label, levels = ord)) |>
  ggplot(aes(y = strata_label)) +
  theme_bw() +
  geom_vline(xintercept = 0, linetype = "dotted") +
  geom_vline(xintercept = gate_meta$scalars$mu, linetype = "dashed",
             colour = "grey40") +
  geom_pointrange(aes(x = gate, xmin = gate_ll, xmax = gate_ul),
                  shape = 21, fill = NA, size = 0.4) +
  geom_point(aes(x = blup), shape = 16, size = 2) +
  labs(x = "E[Y(0) - Y(1) | S], MCS points lost to distress", y = NULL) +
  theme(panel.grid.minor = element_blank())
```

## How much of this rests on one cell?

```{r maihda-loco}
#| output: true
gate_meta$loco |>
  dplyr::arrange(dplyr::desc(pcv_d)) |>
  dplyr::select(dropped_label, tau2_null_d, tau2_main_d, pcv_d, n_zero_d) |>
  knitr::kable(
    digits    = 3,
    col.names = c("Dropped stratum", "tau2 (null)", "tau2 (additive)", "PCV", "n boundary hits"),
    caption   = "PCV recomputed with each stratum dropped in turn, sorted by the resulting PCV. The top row is the stratum whose removal most increases the share attributed to additive main effects -- i.e. the cell carrying the intersectional signal. tau2 and PCV here are pooled arithmetically across imputations, not on the log scale (see the closing discussion): a non-zero 'n boundary hits' means that many of this row's per-imputation refits hit the tau2 = 0 boundary, which is exactly why this row is pooled arithmetically rather than by the headline's log-scale rule."
  )
```

```{r maihda-loco-prose}
#| output: false
# Computed, not hardcoded, so the prose tracks whatever the run produces.
loco_top  <- gate_meta$loco |> dplyr::slice_max(pcv_d, n = 1, with_ties = FALSE)
thin_cell <- gate_meta$cells |> dplyr::slice_min(n_j_min, n = 1, with_ties = FALSE)
# The leverage cell and the thinnest cell are NOT the same cell here, and the
# prose below says so. Do not collapse them into one sentence.
lev_cell  <- gate_meta$cells |>
  dplyr::filter(strata_label == loco_top$dropped_label)
thin_rank <- gate_meta$loco |>
  dplyr::arrange(dplyr::desc(pcv_d)) |>
  dplyr::mutate(rank = dplyr::row_number()) |>
  dplyr::filter(dropped_label == thin_cell$strata_label)
```

Additive sex, race and education account for `r round(100 * gate_meta$scalars$pcv)`% of the
between-stratum variance in the effect of financial distress; the remaining
`r round(100 * (1 - gate_meta$scalars$pcv))`% is intersectional. Three cautions.

First, the decomposition is not evenly supported across cells: dropping
**`r loco_top$dropped_label`** alone moves PCV from `r round(gate_meta$scalars$pcv, 2)` to
`r round(loco_top$pcv_d, 2)`. Note this is *not* the thinnest cell. A random-effects model weights each
stratum by `1/(τ² + SE²)`, so leverage follows precision, not the size of the raw deviation.
**`r loco_top$dropped_label`** carries the result because its departure from the additive surface is
precisely estimated (n ≈ `r lev_cell$n_j_min`, SE `r round(lev_cell$gate_se, 2)`), whereas the thinnest
stratum **`r thin_cell$strata_label`** (n ≈ `r thin_cell$n_j_min`, SE `r round(thin_cell$gate_se, 2)`)
has the largest raw residual but is downweighted to rank `r thin_rank$rank` of 12 in leverage. The
intersectional signal therefore rests on one of the best-estimated strata in the sample, not on a
sparse one.

Second, τ², I² and PCV rest on 12 estimates with 7 residual degrees of freedom; they are descriptive,
and PCV carries no interval. Third, the second stage assumes stratum effects are exchangeable draws
from `N(x_j'β, τ²)` — the same assumption conventional MAIHDA makes, and the price of both τ² and the
shrunken estimates.

Two further notes on interpretation. `I²` and PCV here are **effect-scale analogues** of the MAIHDA
variance partition and proportional change in variance, which are conventionally computed on the
*outcome* scale from a multilevel model of `Y`; they are not the same quantities and should not be
compared numerically against outcome-scale MAIHDA papers.

τ² is pooled two different ways in this section, and deliberately so. The headline decomposition above
(τ²_null, τ²_main, PCV) pools τ² across imputations on the *log* scale, which is a geometric mean: a
boundary estimate of exactly zero in any imputation would pull the pooled value toward zero and PCV
toward one. Of the `r 2 * gate_meta$m` full-model fits behind that headline,
`r gate_meta$n_tau2_zero` hit the τ² = 0 boundary. The leave-one-cell-out table above pools τ²
differently — *arithmetically* — precisely because that layer does reach the boundary in this data: of
its `r 2 * 12 * gate_meta$m` per-imputation refits, `r gate_meta$n_tau2_zero_loco` hit τ² = 0 (see the
table's "n boundary hits" column for which strata). An arithmetic mean handles a boundary value of zero
natively and needs no interval, which matches the LOCO table's role as a descriptive sensitivity check
rather than an inferential estimate; the geometric mean used for the headline would instead have been
dragged toward zero by those same hits. `pool_gate_meta()` warns whenever either layer hits the
boundary.

