---
title: "Full fixed model testing: two waves of data"
author: "Darwin Del Castillo"
format: 
  docx:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from waves 9 and 10, which were collected in 2017-2019 and 2018-2020, respectively. The sample is restricted to individuals with complete data on the outcome and covariates at both time points.

```{r setup}
#| output: false
here::i_am("code/03_full_fixed.qmd")

pacman::p_load(gfoRmula,
               gFormulaMI,
               DiagrammeR,
               tidyverse,
               magrittr,
               colorBlindness,
               ltmle,
               SuperLearner)

source(here::here("fnct", "import_cleaning.R"))
source(here::here("fnct", "helpers.R"))
```

# Model specification: Two waves of data, time-invariant covariates + PCS
In this model, we include two waves of data (waves 9 and 10) and adjust for time-invariant covariates and 
the physical component summary (PCS) score. The model is specified as follows:
**Notation**
- $TI_{cov}$: Time-invariant covariates:
    - Sex
    - Ethnicity
    - Highest educational qualification

```{r model1}
#| label: model1
#| fig-cap: "Estimated mean MCS scores at wave 10 by intervention group and method, adjusting for time-invariant covariates and PCS score."
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=TB]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  C1 [label=<<B><I>PCS<SUB>8</SUB></I></B>>]
                  C2 [label=<<B><I>PCS<SUB>9</SUB></I></B>>]
                  E1 [label=<<B><I>FD<SUB>9</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E2 [label=<<B><I>FD<SUB>10</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y1 [label=<<B><I>MCS<SUB>9</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y2 [label=<<B><I>MCS<SUB>10</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E1 -> E2
                  Y1 -> Y2
                  C1 -> C2

                  TI -> E1, E2, C1, C2, Y1, Y2

                  C2 -> E2
                  C2 -> Y2

                  E1 -> C2

                  C1 -> E1
                  C1 -> Y1

                  E1 -> Y1 
                  E2 -> Y2

                  {rank=min; TI}
                  {rank=same; C1, C2}
                  {rank=same; E1, E2}
                  {rank=max; Y1, Y2}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point.

```{r data_prep}
#| output: false
# Selecting waves 8 (lagged) and 9-10 (scope)
pop_data <- import_data(force = TRUE) |> clean_data() |> preproc_data()

int_data_wide <- pop_data[wave %in% 8:10][, t0 := t0 - 8][
    , pcs_lagged := shift(sf12pcs_dv, type = "lag"), by = pidp
  ][wave %in% 9:10]

# Complete-case dataset
missings <-
  int_data_wide[is.na(sf12mcs_dv) |
             is.na(pcs_lagged) | is.na(sex_dv_base) | is.na(race_base) |
             is.na(econ_dist_bin) | is.na(hiqual_dv_base), .(pidp)]

final_data <- int_data_wide[!missings, on = .(pidp)]

intervention_pattern_2 <- expand_grid(0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()
```


```{r data_prep_long}
# Reshaping to wide format
wide_data <- final_data |>
  mutate(econ_dist_bin = as.factor(econ_dist_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  hiqual_dv_base,
                  race_base),
    outcome = sf12mcs_dv,
    pcs_lagged,
    econ_dist_bin,
    waves = c(0, 1)
  )
```

# Main model: LTMLE with SuperLearner
The main model is estimated using the longitudinal targeted maximum likelihood estimation (LTMLE) method with SuperLearner for model fitting. The SuperLearner library includes a variety of algorithms, such as generalized linear models, random forests, and gradient boosting machines. The results are presented in the figure below, which shows the estimated mean MCS scores at wave 10 by intervention group and method, adjusting for time-invariant covariates and PCS score.

```{r model1}
#| output: false

# Rescaling MCS to [0,1] (Yrange will be c(0,1)); ltmle rescales internally
# for the logistic fluctuation step.

sl_libs <- c("SL.glm", "SL.gam")
MCS_RANGE <- c(0, 1)

# Dedicated frame for ltmle:
#   - rescale Y to [0,1] (don't mutate the shared wide_data)
#   - coerce treatment factor (set for gFormulaImpute) back to 0/1 integer
#   - drop sf12mcs_dv_0 (MCS at wave 9): intermediate outcome, not in this spec
#   - keep columns in W -> A_0 -> L_1 -> A_1 -> Y order
ltmle_data <- wide_data |>
  mutate(
    econ_dist_bin_0 = as.integer(as.character(econ_dist_bin_0)),
    econ_dist_bin_1 = as.integer(as.character(econ_dist_bin_1)),
    sf12mcs_dv_1    = sf12mcs_dv_1 / 100
  ) |>
  select(sex_dv_base,
         hiqual_dv_base,
         race_base,
         pcs_lagged_0,     # W:   PCS at wave 8 (baseline)
         econ_dist_bin_0,  # A_0: FD at wave 9
         pcs_lagged_1,     # L_1: PCS at wave 9 (between A_0 and A_1)
         econ_dist_bin_1,  # A_1: FD at wave 10
         sf12mcs_dv_1)     # Y:   MCS at wave 10

# ltmle's abar accepts a single regime (vector or n x numAnodes matrix for
# dynamic regimes) OR a list of two regimes (for ATE/RR/OR contrasts).
# It does NOT support >2 static regimes in one call, so we re-fit per regime.
# With sl_libs = SL.glm the 4x compute cost is small.
regimes <- list(
  "0-0" = c(0, 0),
  "0-1" = c(0, 1),
  "1-0" = c(1, 0),
  "1-1" = c(1, 1)
)

fit_one_regime <- function(a) {
  ltmle(
    data = ltmle_data,
    Anodes = c("econ_dist_bin_0", "econ_dist_bin_1"),
    Lnodes = "pcs_lagged_1",
    Ynodes = "sf12mcs_dv_1",
    survivalOutcome = FALSE,
    abar = a,
    SL.library = sl_libs,
    estimate.time = FALSE,
    variance.method = "ic",
    Yrange = MCS_RANGE
  )
}

ltmle_fits <- purrr::map(regimes, fit_one_regime)

# For a single-regime ltmle object, summary() exposes the point estimate,
# SE and CI directly under $treatment (no $effect.measures wrapper -- that
# is only present for the 2-regime ltmleEffectMeasures shape).
ltmle_results <- purrr::imap_dfr(ltmle_fits, function(fit, label) {
  tr <- summary(fit, estimator = "tmle")$treatment
  tibble::tibble(
    intervention = label,
    ltmle_effect = tr$estimate,
    ltmle_se     = tr$std.dev,
    ltmle_ll     = tr$CI[1],
    ltmle_ul     = tr$CI[2]
  )
}) |>
  # Y was divided by 100 in data prep; rescale back to 0-100 for comparison.
  dplyr::mutate(dplyr::across(
    c(ltmle_effect, ltmle_se, ltmle_ll, ltmle_ul),
    ~ round(.x * 100, 3)
  ))
```

# Sensitivity models: G-formula-MI, IPTW

## G-formula with multiple imputation of counterfactual outcomes

```{r gformula_mi}
#| output: false
regimes <- dplyr::tibble(a = 0:1, b = 0:1) |>
  expand.grid() |>
  arrange(a, b) |>
  t() |>
  dplyr::as_tibble() |>
  unclass() |> unname()

predictor_matrix <- make_predictor_matrix(wide_data)

# Fixing predictor matrix
## pcs_lagged_0 predicts pcs_lagged_1
predictor_matrix["pcs_lagged_1", "pcs_lagged_0"] <- 1

## econ_dist_bin_0 predicts econ_dist_bin_1
predictor_matrix["econ_dist_bin_1", "econ_dist_bin_0"] <- 1

method_list <- mice::make.method(wide_data)
method_list[c("sex_dv_base", "hiqual_dv_base", "race_base", 
              "pcs_lagged_0", "econ_dist_bin_0", "sf12mcs_dv_0", 
              "pcs_lagged_1", "econ_dist_bin_1", "sf12mcs_dv_1")] <- 
                c("logreg", "pmm", "logreg", 
                  "norm", "logreg", "norm", 
                  "norm", "logreg", "norm")

imps <- gFormulaImpute(
  data = wide_data,
  M = 50,
  trtVars = c("econ_dist_bin_0", "econ_dist_bin_1"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix,
  method = method_list,
  silent = TRUE
)

fits <- imps %$%
  lm(sf12mcs_dv_1 ~ factor(regime) + 0)

outvals <- syntheticPool(fits)

regimes_2 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

## IPTW
```{r iptw}
# Denominator models for IPTW
## P(econ_dist_bin | sex, pcs_lagged, econ_dist_bin)
d0 <- glm(econ_dist_bin_0 ~ sex_dv_base * hiqual_dv_base * race_base * pcs_lagged_0, data = wide_data, family = binomial())
d1 <- glm(econ_dist_bin_1 ~ sex_dv_base * hiqual_dv_base * race_base * pcs_lagged_1 * econ_dist_bin_0, data = wide_data, family = binomial())

# Predicted probabilities
pd0 <- ifelse(wide_data$econ_dist_bin_0 == 1, fitted(d0), 1 - fitted(d0))
pd1 <- ifelse(wide_data$econ_dist_bin_1 == 1, fitted(d1), 1 - fitted(d1))

# Numerators for IPTW
## P(econ_dist_bin | intervention)
n0 <- glm(econ_dist_bin_0 ~ 1, data = wide_data, family = binomial())
n1 <- glm(econ_dist_bin_1 ~ econ_dist_bin_0, data = wide_data, family = binomial())

pn0 <- ifelse(wide_data$econ_dist_bin_0 == 1, fitted(n0), 1 - fitted(n0))
pn1 <- ifelse(wide_data$econ_dist_bin_1 == 1, fitted(n1), 1 - fitted(n1))

# Unstabilized weights
w   <- 1 / (pd0 * pd1)

# Stabilized weights
sw <- (pn0 * pn1) / (pd0 * pd1)

# love plot to check balance after weighting
if (!requireNamespace("cobalt", quietly = TRUE)) {
  install.packages("cobalt")
  library(cobalt)
}
## unstabilized weights
cobalt::bal.tab(
  list(
    econ_dist_bin_0 ~ sex_dv_base + pcs_lagged_0,
    econ_dist_bin_1 ~ sex_dv_base + pcs_lagged_1 + econ_dist_bin_0
  ),
  data = wide_data,
  weights = w
) |>
  cobalt::love.plot(thresholds = c(m = 0.1), stars = "raw") +
  theme_bw() +
  labs(title = "Balance after IPTW weighting") +
  theme(panel.grid.minor = element_blank())

## stabilized weights
cobalt::bal.tab(
  list(
    econ_dist_bin_0 ~ sex_dv_base + pcs_lagged_0,
    econ_dist_bin_1 ~ sex_dv_base + pcs_lagged_1 + econ_dist_bin_0
  ),
  data = wide_data,
  weights = sw
) |>
  cobalt::love.plot(thresholds = c(m = 0.1), stars = "raw") +
  theme_bw() +
  labs(title = "Balance after IPTW weighting") +
  theme(panel.grid.minor = element_blank())

# Outcome model weighted by IPTW
iptw_model <- lm(sf12mcs_dv_1 ~ econ_dist_bin_0 + econ_dist_bin_1, data = wide_data, weights = w)

iptw_model_sw <- lm(sf12mcs_dv_1 ~ econ_dist_bin_0 + econ_dist_bin_1, data = wide_data, weights = sw)

# Summary of IPTW model
summary(iptw_model)
summary(iptw_model_sw)
```

# Results
## Combining results from all methods
```{r results}
# Combine results into a single data frame for plotting

## IPTW results extraction function

a_levels <- levels(wide_data$econ_dist_bin_0)
strategies <- expand_grid(
  econ_dist_bin_0 = factor(a_levels, levels = a_levels),
  econ_dist_bin_1 = factor(a_levels, levels = a_levels)
)

extract_iptw <- function(model, prefix) {
  pred <- predict(model, newdata = strategies, se.fit = TRUE)
  strategies |>
    transmute(
      intervention = paste(as.integer(econ_dist_bin_0) - 1L,
                           as.integer(econ_dist_bin_1) - 1L, sep = "-"),
      "{prefix}_effect" := round(pred$fit, 3),
      "{prefix}_se"     := round(pred$se.fit, 3),
      "{prefix}_ll"     := round(pred$fit - 1.96 * pred$se.fit, 3),
      "{prefix}_ul"     := round(pred$fit + 1.96 * pred$se.fit, 3)
    )
}

iptw_results <- extract_iptw(iptw_model, "iptw_u") |>
  left_join(extract_iptw(iptw_model_sw, "iptw_sw"), by = "intervention") |>
  select("intervention", starts_with("iptw_sw"))


## G-formula MI results
mi_results <- outvals |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_2)

## LTMLE results
ltmle_results <- purrr::imap_dfr(ltmle_fits, function(fit, label) {
  tr <- summary(fit, estimator = "tmle")$treatment
  tibble::tibble(
    intervention = label,
    ltmle_effect = tr$estimate,
    ltmle_se     = tr$std.dev,
    ltmle_ll     = tr$CI[1],
    ltmle_ul     = tr$CI[2]
  )
}) |>
  # Y was divided by 100 in data prep; rescale back to 0-100 for comparison.
  dplyr::mutate(dplyr::across(
    c(ltmle_effect, ltmle_se, ltmle_ll, ltmle_ul),
    ~ round(.x * 100, 3)
  ))

### Merging all results into a single data frame for plotting
comparison <- ltmle_results |>
  left_join(mi_results, by = c("intervention" = "intervention")) |>
  left_join(iptw_results, by = c("intervention" = "intervention")) |>
  relocate(intervention)

tab_models <- comparison |>
  knitr::kable(digits = 3)

```


## Plotting results
```{r plot_results}
# Combine results into a single data frame for plotting
comparison |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("ltmle", "mi", "iptw_sw"),
                    labels = c("LTMLE", "G-Formula MI", "IPTW")),
    intervention = factor(intervention,
                          levels = c("0-0", "1-0", "0-1", "1-1"),
                          labels = c("Never FD", "Early FD", "Late FD", "Always FD"))
  ) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_manual(values = unname(PairedColor12Steps)[7:10],
                      name = "Intervention", guide = "none") +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())
```

