---
title: "Full waves: three waves of data"
author: "Darwin Del Castillo"
format: 
  docx:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from waves 3 and 10, which were collected from 2011 to 2020. The sample is restricted to individuals with complete data on the outcome and covariates at both time points.

```{r setup}
#| output: false
here::i_am("code/04_full_waves.qmd")

pacman::p_load(gfoRmula,
               gFormulaMI,
               DiagrammeR,
               tidyverse,
               magrittr,
               colorBlindness,
               ltmle,
               SuperLearner)

source(here::here("fnct", "import_cleaning.R"))
source(here::here("fnct", "helpers.R"))
```

# Model specification: Two waves of data, time-invariant covariates + PCS
In this model, we include 3 waves of data (waves 3 to 5) and adjust for time-invariant covariates and the physical component summary (PCS) score. The model is specified as follows:
**Notation**
- $TI_{cov}$: Time-invariant covariates:
    - Sex
    - Ethnicity
    - Highest educational qualification

```{r model1}
#| fig-cap: "Estimated mean MCS scores at wave 5, adjusting for time-invariant covariates and PCS score. Nodes are summarised by time point to avoid clutter."
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=TB]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  C1 [label=<<B><I>PCS<SUB>t-1</SUB></I></B>>]
                  C2 [label=<<B><I>PCS<SUB>t</SUB></I></B>>]
                  C3 [label=<<B><I>PCS<SUB>t+1</SUB></I></B>>]
                  E1 [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E2 [label=<<B><I>FD<SUB>t+1</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E3 [label=<<B><I>FD<SUB>t+2</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y2 [label=<<B><I>MCS<SUB>t+1</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y3 [label=<<B><I>MCS<SUB>t+2</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E1 -> E2 -> E3
                  Y1 -> Y2 -> Y3
                  C1 -> C2 -> C3

                  TI -> E1, E2, E3, C1, C2, C3, Y1, Y2, Y3

                  C2 -> E2
                  C2 -> Y2

                  C3 -> E3
                  C3 -> Y3

                  E1 -> C2
                  E2 -> C3

                  C1 -> E1
                  C1 -> Y1

                  E1 -> Y1 
                  E2 -> Y2
                  E3 -> Y3

                  {rank=min; TI}
                  {rank=same; C1, C2, C3}
                  {rank=same; E1, E2, E3}
                  {rank=max; Y1, Y2, Y3}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point.

```{r data_prep}
#| output: false
# Selecting waves 8 (lagged) and 9-10 (scope)
pop_data <- import_data(force = FALSE) |> clean_data() |> preproc_data()

int_data_wide <- pop_data[wave %in% 2:5][, t0 := t0 - 2][
    , `:=` (pcs_lagged = shift(sf12pcs_dv, type = "lag")),
      by = pidp
  ][wave %in% 3:5]

# Complete-case dataset
missings <-
  int_data_wide[is.na(sf12mcs_dv) |
             is.na(pcs_lagged) | is.na(sex_dv_base) | is.na(race_base) |
             is.na(econ_dist_bin) | is.na(hiqual_dv_base), .(pidp)]

final_data <- int_data_wide[!missings, on = .(pidp)]

# We are only interested in always-exposed and never-exposed patterns, so we create a dataset with only those patterns.
intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()
```

```{r data_prep_long}
# Reshaping to wide format
wide_data <- final_data |>
  mutate(econ_dist_bin = as.factor(econ_dist_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  hiqual_dv_base,
                  race_base),
    outcome = sf12mcs_dv,
    pcs_lagged,
    econ_dist_bin,
    waves = c(0, 1, 2)
  )
```

# Main model: LTMLE with SuperLearner
The main model is estimated using the longitudinal targeted maximum likelihood estimation (LTMLE) method with SuperLearner for model fitting. The SuperLearner library includes a variety of algorithms, such as generalized linear models, random forests, and gradient boosting machines. The results are presented in the figure below, which shows the estimated mean MCS scores at wave 10 by intervention group and method, adjusting for time-invariant covariates and PCS score.

```{r ltmle}
#| output: false

# Rescaling MCS to [0,1] (Yrange will be c(0,1)); ltmle rescales internally
# for the logistic fluctuation step.

sl_libs <- c("SL.glm", "SL.gam")
MCS_RANGE <- c(0, 1)

# Dedicated frame for ltmle:
#   - rescale Y to [0,1] (don't mutate the shared wide_data)
#   - coerce treatment factor (set for gFormulaImpute) back to 0/1 integer
#   - drop sf12mcs_dv_0 (MCS at wave 9): intermediate outcome, not in this spec
#   - keep columns in W -> A_0 -> L_1 -> A_1 -> Y order
ltmle_data <- wide_data |>
  mutate(
    econ_dist_bin_0 = as.integer(as.character(econ_dist_bin_0)),
    econ_dist_bin_1 = as.integer(as.character(econ_dist_bin_1)),
    econ_dist_bin_2 = as.integer(as.character(econ_dist_bin_2)),
    sf12mcs_dv_0    = sf12mcs_dv_0 / 100,
    sf12mcs_dv_1    = sf12mcs_dv_1 / 100,
    sf12mcs_dv_2    = sf12mcs_dv_2 / 100
  ) |>
  select(sex_dv_base,
         hiqual_dv_base,
         race_base,
         pcs_lagged_0,     # W:   PCS at wave 2 (baseline)
         econ_dist_bin_0,  # A_0: FD at wave 3
         sf12mcs_dv_0,    # Y_0: MCS at wave 3 (between A_0 and A_1)
         pcs_lagged_1,     # L_1: PCS at wave 3 (between A_0 and A_1)
         econ_dist_bin_1,  # A_1: FD at wave 4
         sf12mcs_dv_1,    # Y_1: MCS at wave 4 (between A_1 and Y)
         pcs_lagged_2,     # L_2: PCS at wave 4 (between A_1 and Y)
         econ_dist_bin_2,  # A_2: FD at wave 5
         sf12mcs_dv_2)     # Y_2:   MCS at wave 5

# ltmle's abar accepts a single regime (vector or n x numAnodes matrix for
# dynamic regimes) OR a list of two regimes (for ATE/RR/OR contrasts).
# It does NOT support >2 static regimes in one call, so we re-fit per regime.
# With sl_libs = SL.glm the 4x compute cost is small.
regimes <- list(
  "0-0-0" = c(0, 0, 0),
  "0-0-1" = c(0, 0, 1),
  "0-1-0" = c(0, 1, 0),
  "0-1-1" = c(0, 1, 1),
  "1-0-0" = c(1, 0, 0),
  "1-0-1" = c(1, 0, 1),
  "1-1-0" = c(1, 1, 0),
  "1-1-1" = c(1, 1, 1)
)

fit_one_regime <- function(a) {
  ltmle(
    data = ltmle_data,
    Anodes = c("econ_dist_bin_0", "econ_dist_bin_1", "econ_dist_bin_2"),
    Lnodes = c("pcs_lagged_0", "pcs_lagged_1", "pcs_lagged_2"),
    Ynodes = c("sf12mcs_dv_0", "sf12mcs_dv_1", "sf12mcs_dv_2"),
    survivalOutcome = FALSE,
    abar = a,
    SL.library = sl_libs,
    estimate.time = FALSE,
    variance.method = "ic",
    Yrange = MCS_RANGE
  )
}

ltmle_fits <- purrr::map(regimes, fit_one_regime)

# For a single-regime ltmle object, summary() exposes the point estimate,
# SE and CI directly under $treatment (no $effect.measures wrapper -- that
# is only present for the 2-regime ltmleEffectMeasures shape).
ltmle_results <- purrr::imap_dfr(ltmle_fits, function(fit, label) {
  tr <- summary(fit, estimator = "tmle")$treatment
  tibble::tibble(
    intervention = label,
    ltmle_effect = tr$estimate,
    ltmle_se     = tr$std.dev,
    ltmle_ll     = tr$CI[1],
    ltmle_ul     = tr$CI[2]
  )
}) |>
  # Y was divided by 100 in data prep; rescale back to 0-100 for comparison.
  dplyr::mutate(dplyr::across(
    c(ltmle_effect, ltmle_se, ltmle_ll, ltmle_ul),
    ~ round(.x * 100, 3)
  ))
```

# Sensitivity analyses: g-formula and IPTW

## G-formula with multiple imputation
```{r gformula-mi}
regimes <- dplyr::tibble(a = 0:1, b = 0:1, c = 0:1) |>
  expand.grid() |>
  arrange(a, b, c) |>
  t() |>
  dplyr::as_tibble(.name_repair = "minimal") |>
  unclass() |> unname()

wide_data_mi <- wide_data |>
  select(-pidp)

predictor_matrix <- make_predictor_matrix(wide_data_mi)

# Fixing predictor matrix
## pcs_lagged_t predicts pcs_lagged_t+1
predictor_matrix["pcs_lagged_1", "pcs_lagged_0"] <- 1
predictor_matrix["pcs_lagged_2", "pcs_lagged_1"] <- 1

## econ_dist_bin_t predicts econ_dist_bin_t+1
predictor_matrix["econ_dist_bin_1", "econ_dist_bin_0"] <- 1
predictor_matrix["econ_dist_bin_2", "econ_dist_bin_1"] <- 1

## sf_12mcs_dv_t only predicts sf_12mcs_dv_t+1 (not other vars)
predictor_matrix[, "sf12mcs_dv_0"] <- 0
predictor_matrix[, "sf12mcs_dv_1"] <- 0
predictor_matrix[, "sf12mcs_dv_2"] <- 0
predictor_matrix["sf12mcs_dv_1", "sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2", "sf12mcs_dv_1"] <- 1
predictor_matrix["regime",] <- 1 # to impute under each regime
predictor_matrix["regime", "regime"] <- 0 # prevent loops

method_list <- mice::make.method(wide_data_mi)
method_list[c("sex_dv_base", "hiqual_dv_base", "race_base", 
              "pcs_lagged_0", "econ_dist_bin_0", "sf12mcs_dv_0", 
              "pcs_lagged_1", "econ_dist_bin_1", "sf12mcs_dv_1",
              "pcs_lagged_2", "econ_dist_bin_2", "sf12mcs_dv_2")] <-
                c("logreg", "pmm", "logreg", 
                  "norm", "logreg", "norm", 
                  "norm", "logreg", "norm",
                  "norm", "logreg", "norm")

imps <- gFormulaImpute(
  data = wide_data_mi,
  M = 50,
  trtVars = c("econ_dist_bin_0", "econ_dist_bin_1", "econ_dist_bin_2"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix,
  method = method_list,
  silent = TRUE
)

fits <- imps %$%
  lm(sf12mcs_dv_2 ~ factor(regime) + 0)

outvals <- syntheticPool(fits)

regimes_3 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```


## MSM - IPTW
```{r iptw}
#| output: false
#| include: false
#| echo: false

# Denominator models for IPTW
## P(A_t | baseline, time-varying confounder, prior treatment)
d0 <- glm(econ_dist_bin_0 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_0,
          data = wide_data, family = binomial())
d1 <- glm(econ_dist_bin_1 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_1 * econ_dist_bin_0,
          data = wide_data, family = binomial())
d2 <- glm(econ_dist_bin_2 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_2 * econ_dist_bin_1,
          data = wide_data, family = binomial())

# Predicted probabilities
pd0 <- ifelse(wide_data$econ_dist_bin_0 == 1, fitted(d0), 1 - fitted(d0))
pd1 <- ifelse(wide_data$econ_dist_bin_1 == 1, fitted(d1), 1 - fitted(d1))
pd2 <- ifelse(wide_data$econ_dist_bin_2 == 1, fitted(d2), 1 - fitted(d2))

# Numerators for IPTW
## P(A_t | prior treatment)
n0 <- glm(econ_dist_bin_0 ~ 1, data = wide_data, family = binomial())
n1 <- glm(econ_dist_bin_1 ~ econ_dist_bin_0, data = wide_data, family = binomial())
n2 <- glm(econ_dist_bin_2 ~ econ_dist_bin_1, data = wide_data, family = binomial())

pn0 <- ifelse(wide_data$econ_dist_bin_0 == 1, fitted(n0), 1 - fitted(n0))
pn1 <- ifelse(wide_data$econ_dist_bin_1 == 1, fitted(n1), 1 - fitted(n1))
pn2 <- ifelse(wide_data$econ_dist_bin_2 == 1, fitted(n2), 1 - fitted(n2))

# Unstabilized weights
w  <- 1 / (pd0 * pd1 * pd2)

# Stabilized weights
sw <- (pn0 * pn1 * pn2) / (pd0 * pd1 * pd2)

# love plot to check balance after weighting
if (!requireNamespace("cobalt", quietly = TRUE)) {
  install.packages("cobalt")
  library(cobalt)
}

## unstabilized weights
w_plot <- cobalt::bal.tab(
  list(
    econ_dist_bin_0 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_0,
    econ_dist_bin_1 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_1 + econ_dist_bin_0 + pcs_lagged_1*econ_dist_bin_0,
    econ_dist_bin_2 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_2 + econ_dist_bin_1 + pcs_lagged_2*econ_dist_bin_1
  ),
  data = wide_data,
  weights = w
) |>
  cobalt::love.plot(thresholds = c(m = 0.1), 
                    var.order = "alphabetical",
                    abs = TRUE,
                    line = FALSE,
                    limits = c(-1.5, 1.5),
                    stars = "raw") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

## stabilized weights
sw_plot <- cobalt::bal.tab(
  list(
    econ_dist_bin_0 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_0,
    econ_dist_bin_1 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_1*econ_dist_bin_0,
    econ_dist_bin_2 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_2*econ_dist_bin_1
  ),
  data = wide_data,
  weights = sw
) |>
  cobalt::love.plot(thresholds = c(m = 0.1), 
                    abs = TRUE,
                    var.order = "alphabetical",
                    line = FALSE,
                    limits = c(-1.5, 1.5),
                    stars = "raw") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

# Outcome model weighted by IPTW
iptw_model <- lm(sf12mcs_dv_2 ~ econ_dist_bin_0 + econ_dist_bin_1 + econ_dist_bin_2 + sex_dv_base + hiqual_dv_base + race_base + econ_dist_bin_0 * econ_dist_bin_1 * econ_dist_bin_2, data = wide_data, weights = w)

iptw_model_sw <- lm(sf12mcs_dv_2 ~ econ_dist_bin_0 + econ_dist_bin_1 + econ_dist_bin_2 + sex_dv_base + hiqual_dv_base + race_base + econ_dist_bin_0 * econ_dist_bin_1 * econ_dist_bin_2, data = wide_data, weights = sw)

# Summary of IPTW model
summary(iptw_model)
summary(iptw_model_sw)
```

```{r iptw_plots}
#| output: true
#| include: true
#| echo: false

# Checking balance after weighting
cowplot::plot_grid(
   w_plot, # unstabilized weights plot
   sw_plot, # stabilized weights plot
   labels = c("A", "B"),
   label_x = 1, hjust = 1.5,
   nrow = 2
 )
```

# Results

## Combining results from all methods
```{r results}
#| output: true
#| include: true
#| echo: false

# Combine results into a single data frame for plotting

## IPTW results extraction function

a_levels <- levels(wide_data$econ_dist_bin_0)
strategies <- expand_grid(
  econ_dist_bin_0 = factor(a_levels, levels = a_levels),
  econ_dist_bin_1 = factor(a_levels, levels = a_levels),
  econ_dist_bin_2 = factor(a_levels, levels = a_levels)
)

extract_iptw <- function(model, prefix) {
  # Outcome model now includes baseline covariates, so the marginal mean
  # under each regime needs g-computation: average the linear predictor
  # across the empirical baseline distribution, rather than predicting at
  # a single synthetic row (which would condition on the reference levels
  # of the factor covariates). SE via delta method on vcov(model); like
  # the prior version this treats IPTW weights as fixed -- bootstrap would
  # be needed to propagate propensity-score uncertainty.
  V   <- vcov(model)
  b   <- coef(model)
  rhs <- delete.response(terms(model))

  build_contrast <- function(a0, a1, a2) {
    cf <- wide_data |>
      dplyr::mutate(
        econ_dist_bin_0 = factor(as.character(a0), levels = a_levels),
        econ_dist_bin_1 = factor(as.character(a1), levels = a_levels),
        econ_dist_bin_2 = factor(as.character(a2), levels = a_levels)
      )
    colMeans(model.matrix(rhs, data = cf))
  }

  L <- do.call(rbind, purrr::pmap(
    list(strategies$econ_dist_bin_0,
         strategies$econ_dist_bin_1,
         strategies$econ_dist_bin_2),
    build_contrast
  ))

  est <- as.numeric(L %*% b)
  se  <- sqrt(diag(L %*% V %*% t(L)))

  strategies |>
    transmute(
      intervention = paste(as.integer(econ_dist_bin_0) - 1L,
                           as.integer(econ_dist_bin_1) - 1L,
                           as.integer(econ_dist_bin_2) - 1L, sep = "-"),
      "{prefix}_effect" := round(est, 3),
      "{prefix}_se"     := round(se, 3),
      "{prefix}_ll"     := round(est - 1.96 * se, 3),
      "{prefix}_ul"     := round(est + 1.96 * se, 3)
    )
}

iptw_results <- extract_iptw(iptw_model, "iptw_u") |>
  left_join(extract_iptw(iptw_model_sw, "iptw_sw"), by = "intervention") |>
  select("intervention", starts_with("iptw_sw"))


## G-formula MI results
mi_results <- outvals |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_3)

## LTMLE results
ltmle_results <- purrr::imap_dfr(ltmle_fits, function(fit, label) {
  tr <- summary(fit, estimator = "tmle")$treatment
  tibble::tibble(
    intervention = label,
    ltmle_effect = tr$estimate,
    ltmle_se     = tr$std.dev,
    ltmle_ll     = tr$CI[1],
    ltmle_ul     = tr$CI[2]
  )
}) |>
  # Y was divided by 100 in data prep; rescale back to 0-100 for comparison.
  dplyr::mutate(dplyr::across(
    c(ltmle_effect, ltmle_se, ltmle_ll, ltmle_ul),
    ~ round(.x * 100, 3)
  ))

### Merging all results into a single data frame for plotting
comparison <- ltmle_results |>
  left_join(mi_results, by = c("intervention" = "intervention")) |>
  left_join(iptw_results, by = c("intervention" = "intervention")) |>
  relocate(intervention)

tab_models <- comparison |>
  knitr::kable(digits = 3)

print(tab_models)
```

## Plotting results
```{r plot_results}
#| output: true
#| include: true

# Combine results into a single data frame for plotting
comparison |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("ltmle", "mi", "iptw_sw"),
                    labels = c("LTMLE", "G-Formula MI", "IPTW")),
    intervention = factor(intervention,
                          levels = c("0-0-0", "0-0-1", "0-1-0", "0-1-1", "1-0-0", "1-0-1", "1-1-0", "1-1-1"),
  )) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Intervention",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(39, 51, by = 2)) +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())
```

