---
title: "New pipelines: multiple imputation"
author: "Darwin Del Castillo"
format: 
  docx:
    number-sections: true
    
editor: source
---

# Data
Data comes from the UK Household Longitudinal Study (UKHLS), also known as Understanding Society. We use data from waves 3 and 10, which were collected from 2011 to 2020. The sample is restricted to individuals with complete data on the outcome and covariates at both time points.

```{r setup}
#| output: false
here::i_am("code/05_imputation.qmd")

pacman::p_load(gfoRmula,
               gFormulaMI,
               DiagrammeR,
               tidyverse,
               magrittr,
               colorBlindness,
               ltmle,
               mice,
               SuperLearner)

source(here::here("fnct", "import_cleaning.R"))
source(here::here("fnct", "helpers.R"))
```

# Model specification: Two waves of data, time-invariant covariates + PCS
In this model, we include 3 waves of data (waves 3 to 5) and adjust for time-invariant covariates and the physical component summary (PCS) score. The model is specified as follows:
**Notation**
- $TI_{cov}$: Time-invariant covariates:
    - Sex
    - Ethnicity
    - Highest educational qualification

```{r model1}
#| label: model1
#| fig-cap: "Estimated mean MCS scores at wave 5, adjusting for time-invariant covariates and PCS score. Nodes are summarised by time point to avoid clutter."
#| fig-width: 8
#| fig-height: 10
#| fig-align: center

DiagrammeR::grViz("
                  digraph{
                  graph[rankdir=TB]

                  node[fontname=Arial]
                  TI [label=<<B><I>TI<SUB>cov</SUB></I></B>>]
                  C1 [label=<<B><I>PCS<SUB>t-1</SUB></I></B>>]
                  C2 [label=<<B><I>PCS<SUB>t</SUB></I></B>>]
                  C3 [label=<<B><I>PCS<SUB>t+1</SUB></I></B>>]
                  E1 [label=<<B><I>FD<SUB>t</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E2 [label=<<B><I>FD<SUB>t+1</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  E3 [label=<<B><I>FD<SUB>t+2</SUB></I></B>>, 
                      fontcolor=deepskyblue4, 
                      color=deepskyblue4]
                  Y1 [label=<<B><I>MCS<SUB>t</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y2 [label=<<B><I>MCS<SUB>t+1</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]
                  Y3 [label=<<B><I>MCS<SUB>t+2</SUB></I></B>>, 
                      fontcolor=darkorange2, 
                      color=darkorange2]

                  E1 -> E2 -> E3
                  Y1 -> Y2 -> Y3
                  C1 -> C2 -> C3

                  TI -> E1, E2, E3, C1, C2, C3, Y1, Y2, Y3

                  C2 -> E2
                  C2 -> Y2

                  C3 -> E3
                  C3 -> Y3

                  E1 -> C2
                  E2 -> C3

                  C1 -> E1
                  C1 -> Y1

                  E1 -> Y1 
                  E2 -> Y2
                  E3 -> Y3

                  {rank=min; TI}
                  {rank=same; C1, C2, C3}
                  {rank=same; E1, E2, E3}
                  {rank=max; Y1, Y2, Y3}
            }
                  "
)
```

# Data preparation
The data is prepared by creating a long format dataset with one row per individual per time point.

```{r data_prep}
#| output: false
# Selecting waves 8 (lagged) and 9-10 (scope)
pop_data <- import_data(force = FALSE) |> clean_data() |> preproc_data()

int_data_wide <- pop_data[wave %in% 2:5][, t0 := t0 - 2][
    , `:=` (pcs_lagged = shift(sf12pcs_dv, type = "lag")),
      by = pidp
  ][wave %in% 3:5]

# no missing data exclusion
missings <-
  int_data_wide[is.na(sf12mcs_dv) |
             is.na(pcs_lagged) | is.na(sex_dv_base) | is.na(race_base) |
             is.na(econ_dist_bin) | is.na(hiqual_dv_base), .(pidp)]

# final_data <- int_data_wide[!missings, on = .(pidp)]

final_data <- int_data_wide

# We are only interested in always-exposed and never-exposed patterns, so we create a dataset with only those patterns.
intervention_pattern_3 <- expand_grid(0:1, 0:1, 0:1) |>
  t() |>
  as_tibble(.name_repair = "minimal") |>
  unclass()

# Seed
seed_random <- 20260522
```

```{r data_prep_wide}
# Reshaping to wide format
wide_data <- final_data |>
  mutate(econ_dist_bin = as.factor(econ_dist_bin)) |>
  make_wide(
    pidp,
    t0,
    base_cols = c(sex_dv_base,
                  hiqual_dv_base,
                  race_base),
    outcome = sf12mcs_dv,
    pcs_lagged,
    econ_dist_bin,
    waves = c(0, 1, 2)
  )
```

# Multiple imputation (wide format)
```{r mice-gformula}
wide_data_mi <- wide_data |>
  select(-pidp)

method_list <- mice::make.method(wide_data_mi)
method_list[c("sex_dv_base", "hiqual_dv_base", "race_base", 
              "pcs_lagged_0", "econ_dist_bin_0", "sf12mcs_dv_0", 
              "pcs_lagged_1", "econ_dist_bin_1", "sf12mcs_dv_1",
              "pcs_lagged_2", "econ_dist_bin_2", "sf12mcs_dv_2")] <-
                c("", "polr", "logreg", 
                  "pmm", "logreg", "pmm", 
                  "pmm", "logreg", "pmm",
                  "pmm", "logreg", "pmm")

# predictor matrix for mice imputation
wide_mids <- mice(data = wide_data_mi,
                  defaultMethod = c("pmm", "logreg", "polyreg", "polr"), 
                  m = 35, 
                  maxit = 15,
                  seed = seed_random,
                  method = method_list)
```

## Additional: mice diagnostics
```{r mice-diag}
#| fig-width: 12
#| fig-height: 14
#| fig-align: center

# Diagnostic plots for mice imputation
bwplot(wide_mids)

## Checking goodness of fitness
fit <- with(wide_mids, glm(ici(wide_mids) ~ sf12mcs_dv_0 + pcs_lagged_0 + econ_dist_bin_0 + sf12mcs_dv_1 + pcs_lagged_1 + econ_dist_bin_1 + sf12mcs_dv_2 + pcs_lagged_2 + econ_dist_bin_2 + hiqual_dv_base + race_base + sex_dv_base,
                          family = binomial))

ps <- rep(rowMeans(sapply(fit$analyses, fitted.values)),
          wide_mids$m + 1)

numeric_vars <- colnames(wide_data_mi)[
  vapply(wide_data_mi, is.numeric, logical(1))
]

fitted_plot <- xyplot(
  wide_mids,
  as.formula(paste(paste(numeric_vars, collapse = " + "),
                   "~ ps | as.factor(.imp)")),
  outer    = TRUE,
  as.table = TRUE,
  xlab     = "Probability that record is incomplete",
  ylab     = "Value",
  pch      = c(1, 19),
  cex      = 1.1,
  col      = mdc(1:2),
  scales   = list(
    y = list(relation = "free", cex = 0.7),
    x = list(cex = 0.7)
  ),
  par.strip.text = list(cex = 0.7)
)

fitted_plot

# Convergence plot
plot(wide_mids)

# Worm plot for one outcome
wp_mids <- function(mids_obj, var, level = 0.95) {
  long <- mice::complete(mids_obj, action = "long", include = TRUE)
  long$.imp <- factor(long$.imp,
                      levels = 0:mids_obj$m,
                      labels = c("Observed",
                                 paste("Imp.", seq_len(mids_obj$m))))
  long$.z <- as.numeric(stats::ave(long[[var]], long$.imp,
                                   FUN = function(v) scale(v)[, 1]))
  long <- long[!is.na(long$.z), ]

  crit <- qnorm((1 + level) / 2)

  lattice::qqmath(
    ~ .z | .imp, data = long,
    aspect   = 1,
    as.table = TRUE,
    xlim     = c(-3, 3),
    ylim     = c(-0.6, 0.6),
    xlab     = "Unit normal quantile",
    ylab     = "Deviation",
    main     = sprintf(var),
    par.strip.text = list(cex = 0.7),
    panel = function(x, ...) {
      qq    <- stats::qqnorm(x, plot.it = FALSE)
      detr  <- qq$y - qq$x
      n     <- length(qq$y)
      z_seq <- seq(-2.75, 2.75, by = 0.05)
      p     <- pnorm(z_seq)
      se    <- sqrt(p * (1 - p) / n) / dnorm(z_seq)

      lattice::panel.abline(h = 0, lty = 2, col = "grey60")
      lattice::panel.lines(z_seq,  crit * se, lty = 2, col = "grey40")
      lattice::panel.lines(z_seq, -crit * se, lty = 2, col = "grey40")
      lattice::panel.points(qq$x, detr,
                            pch = 19, cex = 0.4, col = mice::mdc(2))
    }
  )
}

wp_mids(wide_mids, "sf12mcs_dv_0")
wp_mids(wide_mids, "sf12mcs_dv_1")
wp_mids(wide_mids, "sf12mcs_dv_2")
wp_mids(wide_mids, "pcs_lagged_0")
wp_mids(wide_mids, "pcs_lagged_1")
wp_mids(wide_mids, "pcs_lagged_2")

# density plots for imputed vs observed data
densityplot(wide_mids, ~ sf12mcs_dv_0)
densityplot(wide_mids, ~ sf12mcs_dv_1)
densityplot(wide_mids, ~ sf12mcs_dv_2)

densityplot(wide_mids, ~ pcs_lagged_0)
densityplot(wide_mids, ~ pcs_lagged_1)
densityplot(wide_mids, ~ pcs_lagged_2)

densityplot(wide_mids, ~ econ_dist_bin_0)
densityplot(wide_mids, ~ econ_dist_bin_1)
densityplot(wide_mids, ~ econ_dist_bin_2)

densityplot(wide_mids, ~ hiqual_dv_base)
```

# `mice` positivity check
```{r mice-positivity}
positivity_check <- function(mids_obj, data_list) {
    map(data_list, \(dat) {
        dat |>
            group_by(econ_dist_bin_0,
                     econ_dist_bin_1,
                     econ_dist_bin_2) |>
            summarise(n = n(), .groups = "drop")
    })
}

positivity_results <- positivity_check(wide_mids, 
                                       mice::complete(wide_mids, action = "all"))

print(positivity_results)
```

# Main analysis: LTMLE with multiple imputation
```{r ltmle-mi}
#| output: false
#| include: true

# Rescaling MCS to [0,1] (Yrange will be c(0,1)); ltmle rescales internally
# for the logistic fluctuation step.

sl_libs <- c("SL.glm", "SL.gam")
MCS_RANGE <- c(0, 1)

# Reordering each imputed dataset to match the ltmle spec:

ltmle_prep <- function(mira_obj, data_list) {
    map(data_list, \(dat) {
      dat |>
        mutate(
          econ_dist_bin_0 = as.integer(as.character(econ_dist_bin_0)),
          econ_dist_bin_1 = as.integer(as.character(econ_dist_bin_1)),
          econ_dist_bin_2 = as.integer(as.character(econ_dist_bin_2)),
          sf12mcs_dv_0    = sf12mcs_dv_0 / 100,
          sf12mcs_dv_1    = sf12mcs_dv_1 / 100,
          sf12mcs_dv_2    = sf12mcs_dv_2 / 100
        ) |>
        select (sex_dv_base,
                hiqual_dv_base,
                race_base,
                pcs_lagged_0,     # W:   PCS at wave 2 (baseline)
                econ_dist_bin_0,  # A_0: FD at wave 3
                sf12mcs_dv_0,    # Y_0: MCS at wave 3 (between A_0 and A_1)
                pcs_lagged_1,     # L_1: PCS at wave 3 (between A_0 and A_1)
                econ_dist_bin_1,  # A_1: FD at wave 4
                sf12mcs_dv_1,    # Y_1: MCS at wave 4 (between A_1 and Y)
                pcs_lagged_2,     # L_2: PCS at wave 4 (between A_1 and Y)
                econ_dist_bin_2,  # A_2: FD at wave 5
                sf12mcs_dv_2)     # Y_2:   MCS at wave 5
    })
}

ltmle_data_list <- ltmle_prep(wide_mids, 
                              mice::complete(wide_mids, action = "all"))

regimes <- list(
  "0-0-0" = c(0, 0, 0),
  "0-0-1" = c(0, 0, 1),
  "0-1-0" = c(0, 1, 0),
  "0-1-1" = c(0, 1, 1),
  "1-0-0" = c(1, 0, 0),
  "1-0-1" = c(1, 0, 1),
  "1-1-0" = c(1, 1, 0),
  "1-1-1" = c(1, 1, 1)
)

Qform <- c(
    pcs_lagged_0 = "Q.kplus1 ~ race_base + sex_dv_base + hiqual_dv_base",
    pcs_lagged_1 = "Q.kplus1 ~ race_base + sex_dv_base + hiqual_dv_base + econ_dist_bin_0 + pcs_lagged_0",
    pcs_lagged_2 = "Q.kplus1 ~ race_base + sex_dv_base + hiqual_dv_base +
                               econ_dist_bin_0 + pcs_lagged_0 + pcs_lagged_1 +
                               econ_dist_bin_1",
    sf12mcs_dv_0    = "Q.kplus1 ~ race_base + sex_dv_base + hiqual_dv_base + pcs_lagged_0 + econ_dist_bin_0",
    sf12mcs_dv_1    = "Q.kplus1 ~ race_base + sex_dv_base + hiqual_dv_base + pcs_lagged_0 + econ_dist_bin_0 + pcs_lagged_1 + econ_dist_bin_1 + sf12mcs_dv_0",
    sf12mcs_dv_2    = "Q.kplus1 ~ race_base + sex_dv_base + hiqual_dv_base + pcs_lagged_0 + econ_dist_bin_0 + pcs_lagged_1 + econ_dist_bin_1 + pcs_lagged_2 + econ_dist_bin_2 + sf12mcs_dv_0 + sf12mcs_dv_1"
)

gform <- c(
    econ_dist_bin_0 = "econ_dist_bin_0 ~ race_base + sex_dv_base + hiqual_dv_base + pcs_lagged_0",
    econ_dist_bin_1 = "econ_dist_bin_1 ~ race_base + sex_dv_base + hiqual_dv_base + pcs_lagged_1 + econ_dist_bin_0",
    econ_dist_bin_2 = "econ_dist_bin_2 ~ race_base + sex_dv_base + hiqual_dv_base + pcs_lagged_2 + econ_dist_bin_0 + econ_dist_bin_1"
)

# this function will run ltmle for one regime across each imputed dataset, returning a list of ltmle fits that can be pooled with mice::pool()

fit_one_regime_per_imp <- function(regime_value, data) {
    map(data, \(dat) {
            ltmle(
                data = dat,
                Anodes = c("econ_dist_bin_0", "econ_dist_bin_1", "econ_dist_bin_2"),
                Lnodes = c("pcs_lagged_0", "pcs_lagged_1", "pcs_lagged_2"),
                Ynodes = c("sf12mcs_dv_0", "sf12mcs_dv_1", "sf12mcs_dv_2"),
                survivalOutcome = FALSE,
                Qform = Qform,
                gform = gform,
                abar = regime_value,
                SL.library = sl_libs,
                estimate.time = FALSE,
                variance.method = "ic",
                Yrange = MCS_RANGE
            )
    })
}

ltmle_fits <- purrr::map(regimes, fit_one_regime_per_imp, data = ltmle_data_list)

# ltmle_fits[[label]] is now a list of m single-regime fits (one per
# imputation). summary(fit, estimator = "tmle")$treatment exposes the
# influence-curve point estimate and SE on each one; we then pool the
# scalar estimand across imputations with Rubin's rules via
# mice::pool.scalar (T = U_bar + (1 + 1/m) * B).
ltmle_results <- purrr::imap_dfr(ltmle_fits, function(fits_per_imp, label) {
  trs <- purrr::map(fits_per_imp,
                    \(fit) summary(fit, estimator = "tmle")$treatment)
  Q   <- purrr::map_dbl(trs, "estimate")
  U   <- purrr::map_dbl(trs, "std.dev")^2
  pooled <- mice::pool.scalar(Q = Q, U = U, n = Inf, k = 1)
  se  <- sqrt(pooled$t)
  tibble::tibble(
    intervention = label,
    ltmle_effect = pooled$qbar,
    ltmle_se     = se,
    ltmle_ll     = pooled$qbar - 1.96 * se,
    ltmle_ul     = pooled$qbar + 1.96 * se
  )
}) |>
  # Y was divided by 100 in data prep; rescale back to 0-100 for comparison.
  dplyr::mutate(dplyr::across(
    c(ltmle_effect, ltmle_se, ltmle_ll, ltmle_ul),
    ~ round(.x * 100, 3)
  ))
```


# Sensitivity analyses: g-formula and IPTW

## G-formula via multiple imputation

### Imputing counterfactual outcomes under each regime using `gFormulaMI`
```{r gformula-mi}
regimes <- dplyr::tibble(a = 0:1, b = 0:1, c = 0:1) |>
  expand.grid() |>
  arrange(a, b, c) |>
  t() |>
  dplyr::as_tibble(.name_repair = "minimal") |>
  unclass() |> unname()

predictor_matrix <- make_predictor_matrix(wide_data_mi)

# Fixing predictor matrix
## pcs_lagged_t predicts pcs_lagged_t+1
predictor_matrix["pcs_lagged_1", "pcs_lagged_0"] <- 1
predictor_matrix["pcs_lagged_2", "pcs_lagged_1"] <- 1

## econ_dist_bin_t predicts econ_dist_bin_t+1
predictor_matrix["econ_dist_bin_1", "econ_dist_bin_0"] <- 1
predictor_matrix["econ_dist_bin_2", "econ_dist_bin_1"] <- 1

## sf_12mcs_dv_t only predicts sf_12mcs_dv_t+1 (not other vars)
predictor_matrix[, "sf12mcs_dv_0"] <- 0
predictor_matrix[, "sf12mcs_dv_1"] <- 0
predictor_matrix[, "sf12mcs_dv_2"] <- 0
predictor_matrix["sf12mcs_dv_1", "sf12mcs_dv_0"] <- 1
predictor_matrix["sf12mcs_dv_2", "sf12mcs_dv_1"] <- 1
predictor_matrix["regime",] <- 1 # to impute under each regime
predictor_matrix["regime", "regime"] <- 0 # prevent loops

imps <- gFormulaImpute(
  data = wide_mids,
  M = 50,
  trtVars = c("econ_dist_bin_0", "econ_dist_bin_1", "econ_dist_bin_2"),
  trtRegimes = regimes,
  predictorMatrix = predictor_matrix,
#  method = method_list, # gformula inherits info from mice object
  silent = TRUE
)

fits <- imps %$%
  lm(sf12mcs_dv_2 ~ factor(regime) + 0)

outvals <- syntheticPool(fits)

regimes_3 <- tibble(intervention = regimes |> 
  map(paste, collapse = "-") |> 
  reduce(c)
)
```

## IPTW with multiple imputation

```{r iptw-mi}
# As the mids object contains different imputed datasets, we need to use each dataset for the IPTW calculation.

# Fit denominator models, build per-imputation IPTW weights, and run the
# weighted outcome model inside one with() so everything stays aligned with
# wide_mids. Result is a mira ready for mice::pool().
iptw_fit <- with(wide_mids, {
  # Denominators: P(A_t | baseline, time-varying confounder, prior treatment)
  d0 <- glm(econ_dist_bin_0 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_0,
            family = binomial())
  d1 <- glm(econ_dist_bin_1 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_1 * econ_dist_bin_0,
            family = binomial())
  d2 <- glm(econ_dist_bin_2 ~ sex_dv_base + hiqual_dv_base + race_base + pcs_lagged_2 * econ_dist_bin_1,
            family = binomial())

  # Numerators: P(A_t | prior treatment) — stabilising marginals
  n0 <- glm(econ_dist_bin_0 ~ 1, family = binomial())
  n1 <- glm(econ_dist_bin_1 ~ econ_dist_bin_0, family = binomial())
  n2 <- glm(econ_dist_bin_2 ~ econ_dist_bin_1, family = binomial())

  sw <- reduce(pmap(list(list(n0, n1, n2),
                         list(d0, d1, d2),
                         list(econ_dist_bin_0, econ_dist_bin_1, econ_dist_bin_2)),
                    \(n, d, a) ifelse(a == 1,
                                      fitted(n) / fitted(d),
                                      (1 - fitted(n)) / (1 - fitted(d)))),
               `*`)

  lm(sf12mcs_dv_2 ~ econ_dist_bin_0 + econ_dist_bin_1 + econ_dist_bin_2 +
       sex_dv_base + hiqual_dv_base + race_base +
       econ_dist_bin_0 * econ_dist_bin_1 * econ_dist_bin_2,
     weights = sw)
})

iptw_pool <- mice::pool(iptw_fit)
summary(iptw_pool)
```

### Table with results from all models
```{r table-results}
## IPTW results
## IPTW results extraction function

a_levels <- levels(wide_data$econ_dist_bin_0)
strategies <- expand_grid(
  econ_dist_bin_0 = factor(a_levels, levels = a_levels),
  econ_dist_bin_1 = factor(a_levels, levels = a_levels),
  econ_dist_bin_2 = factor(a_levels, levels = a_levels)
)

extract_iptw <- function(mira_obj, data_list, prefix) {
  # G-computation marginal means from the IPTW-weighted lm, pooled across
  # imputations. The contrast vector L for each regime is averaged across
  # imputations (baseline covariate distribution differs slightly between
  # completed datasets) before contracting with the pooled coefficient
  # vector. Between-imputation variance is propagated via Rubin's rules:
  # T = W_bar + (1 + 1/m) * B. Propensity-score uncertainty within each
  # imputation is still treated as fixed -- bootstrap within each
  # imputation would be needed to propagate that too.
  m     <- length(mira_obj$analyses)
  coefs <- map(mira_obj$analyses, coef)
  vcovs <- map(mira_obj$analyses, vcov)
  b     <- reduce(coefs, `+`) / m
  W     <- reduce(vcovs, `+`) / m
  B     <- map(coefs, \(bk) tcrossprod(bk - b)) |> reduce(`+`) / (m - 1)
  V     <- W + (1 + 1 / m) * B
  rhs   <- delete.response(terms(mira_obj$analyses[[1]]))

  build_contrast <- function(a0, a1, a2) {
    map(data_list, \(dat) {
      cf <- dat |>
        dplyr::mutate(
          econ_dist_bin_0 = factor(as.character(a0), levels = a_levels),
          econ_dist_bin_1 = factor(as.character(a1), levels = a_levels),
          econ_dist_bin_2 = factor(as.character(a2), levels = a_levels)
        )
      colMeans(model.matrix(rhs, data = cf))
    }) |>
      reduce(`+`) / length(data_list)
  }

  L <- do.call(rbind, pmap(
    list(strategies$econ_dist_bin_0,
         strategies$econ_dist_bin_1,
         strategies$econ_dist_bin_2),
    build_contrast
  ))

  est <- as.numeric(L %*% b)
  se  <- sqrt(diag(L %*% V %*% t(L)))

  strategies |>
    transmute(
      intervention = paste(as.integer(econ_dist_bin_0) - 1L,
                           as.integer(econ_dist_bin_1) - 1L,
                           as.integer(econ_dist_bin_2) - 1L, sep = "-"),
      "{prefix}_effect" := round(est, 3),
      "{prefix}_se"     := round(se, 3),
      "{prefix}_ll"     := round(est - 1.96 * se, 3),
      "{prefix}_ul"     := round(est + 1.96 * se, 3)
    )
}

iptw_results <- extract_iptw(
  mira_obj  = iptw_fit,
  data_list = mice::complete(wide_mids, action = "all"),
  prefix    = "iptw_sw"
) |>
  select("intervention", starts_with("iptw_sw"))

## G-formula MI results
mi_results <- outvals |>
  as_tibble() |>
  rownames_to_column("Intervention") |>
  transmute(mi_effect = Estimate,
            mi_se = sqrt(Total),
            mi_ll = `95% CI L`,
            mi_ul = `95% CI U`) |>
  bind_cols(regimes_3)

comparison <- ltmle_results |>
  left_join(mi_results, by = "intervention") |>
  left_join(iptw_results, by = "intervention") |>
  relocate("intervention")
```

```{r plot_results}
#| output: true
#| include: true

# Combine results into a single data frame for plotting
comparison |>
  pivot_longer(
    -intervention,
    names_to = c("method", ".value"),
    names_pattern = "(.*)_(.*)"
  ) |>
  mutate(
    method = factor(method,
                    levels = c("ltmle", "mi", "iptw_sw"),
                    labels = c("LTMLE", "G-Formula MI", "IPTW")),
    intervention = factor(intervention,
                          levels = c("0-0-0", "0-0-1", "0-1-0", "0-1-1", "1-0-0", "1-0-1", "1-1-0", "1-1-1"),
  )) |>
  ggplot(aes(x = effect, y = intervention,
             colour = intervention,
             xmin = ll, xmax = ul)) +
  theme_bw() +
  geom_point(size = 2.5) +
  geom_errorbar(orientation = "y", width = 0.3) +
  scale_colour_brewer(palette = "Set1",
                      name = "Intervention",
                      guide = "legend") +
  scale_x_continuous(breaks = seq(39, 51, by = 2)) +
  labs(x = "Mean MCS score (95% CI)", y = NULL) +
  facet_wrap(~method, ncol = 1) +
  theme(panel.grid.minor = element_blank())
```